---
title: "MRMhub Data Processing Workflow for Dataset 4"
subtitle: "Supplementary Code 3"
author:
  - name: "Bo Burla"
    affiliation: "National University of Singapore"
  - name: "Guo Shou Teo"
    affiliation: "National University of Singapore"
  - name: "Hyungwon Choi"
    affiliation: "National University of Singapore"
date: "`r Sys.Date()`"
format:
  html:
    code-tools: true
    self-contained: true
    code-overflow: wrap
    toc: true
    fig-align: center
  pdf:
    pdf-engine: lualatex
    fig-format: pdf
    toc: true
    fig-dpi: 300
    number-sections: true
    keep-tex: false
    include-in-header: caption-font.tex
bibliography: references.bib
---

```{r chunk1-init, include = FALSE, echo = FALSE}
#| cache: false
# Rendering options for this document. Caching is off: the dataset is small
# (15 analytes, ~31 analyses) so a full render is fast, and several chunks write
# intermediate files, where stale cache state would be a liability.
knitr::opts_chunk$set(cache = FALSE, collapse = TRUE, results = TRUE, message = TRUE, warning = TRUE, comment = "#>", out.width = "100%")
set.seed(1041)
options(dplyr.print_max = 15)

library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(ggplot2)
library(patchwork)

# Data root for this notebook (flat layout, mirroring Dataset 1/3).
d4 <- "./data/dataset-4"
dir.create("output", showWarnings = FALSE, recursive = TRUE)
```

\newpage

# Overview

Dataset 4 is a **fully quantitative steroid panel** used to validate MRMhub
against a vendor pipeline (Agilent MassHunter). Fifteen steroids were measured in
human serum with the IBL International *Steroid Panel LC-MS* kit
(cat. 30191875) according to the instructions for use, with modifications. The
kit acquires its analytes over two chromatographic panels; **only Panel 1 (15
analytes) is used here**, which is why every sample identifier carries the `_P1`
suffix (e.g. `Cal A_P1`, `QC_High1_P1`).

Unlike Datasets 1 and 3 (relative lipidomics), this is an externally calibrated
absolute assay. The analytical design comprises an **external calibration curve**
(`Cal 0`–`Cal F`), **low and high quality-control samples** (`QC_Low1`,
`QC_High1`), and **external quality-assessment (EQA) samples** — SKML ring-trial
materials with reference-method-assigned target concentrations [@jansen2014].

The workflow compares MRMhub and MassHunter at both pipeline stages:

- **INTEGRATOR** — peak areas from MRMhub-INTEGRATOR vs a MassHunter
  re-integration of the same raw files (Figure 1).
- **QUANT** — calibration fits and concentrations, ending in accuracy against the
  SKML EQA reference values (Figure 2).

# Raw Data Processing: Peak Picking and Integration

Raw `.d` files were converted to mzML and integrated with MRMhub-INTEGRATOR, as
described for Datasets 1 and 3. The batch was integrated in two subsets — the
**system-suitability** injections (`SST`, ten replicate injections) and the
**assay** samples (`ASSAY`: blanks, calibrators, QCs, EQA samples) — exported as
separate long-format tables and combined downstream. INTEGRATOR reports peak
areas, retention times and peak metadata in `long.csv`, and writes per-transition
integration plots (the shaded, integrated region) to the `by_transition` folder.

The same raw files were independently re-integrated in Agilent MassHunter
Quantitative Analysis, providing the reference peak areas and concentrations used
throughout this document.

```{r chunk2-integrator-examples}
#| fig-width: 8
#| fig-height: 3.2
#| fig-cap: "Example MRMhub-INTEGRATOR chromatograms (Panel 1 quantifier transitions, high QC sample). The shaded region is the integrated peak."
#| fig.alt: "Two example integrated chromatograms for Cortisol and Testosterone."
# Chromatograms are pre-cropped from the INTEGRATOR by_transition PDFs by
# figures/dataset4/dataset4-chromatograms.R and embedded here as static PNGs, so
# the notebook render needs neither the PDFs nor magick/ghostscript.
chrom_dir  <- "images/dataset4"
chrom_grob <- function(name) grid::rasterGrob(
  png::readPNG(file.path(chrom_dir, paste0(name, ".png"))), interpolate = TRUE)

chrom_gg <- function(name, title, subtitle = NULL) {
  ggplot() + annotation_custom(chrom_grob(name)) +
    labs(title = title, subtitle = subtitle) +
    theme_void(base_size = 8, base_family = "sans") +
    theme(plot.title    = element_text(size = 8,   hjust = 0.5, margin = margin(b = 1)),
          plot.subtitle = element_text(size = 6.8, hjust = 0.5, colour = "grey30",
                                       margin = margin(b = 2)))
}

(chrom_gg("integrator_Cortisol",     "Cortisol") |
 chrom_gg("integrator_Testosterone", "Testosterone"))
```

::: {.callout-note}
The example chromatograms are cropped from the INTEGRATOR `by_transition` PDFs by
`figures/dataset4/dataset4-chromatograms.R` (run once; requires the `magick` R
package and **ghostscript**) and embedded here as static images from
`images/dataset4/`. The notebook render itself needs neither.
:::

# Data Postprocessing and QC

The postprocessing pipeline is the QUANT module (`library(mrmhub)`): import the
INTEGRATOR results and metadata, normalize to internal standards, fit external
calibration curves, and quantify. This is a shorter pipeline than Datasets 1/3 —
no drift/batch correction or feature filtering, which do not apply to a
15-analyte, single-batch panel.

```{r chunk3-loadpkg}
#| cache: false
library(mrmhub)
```

## Import INTEGRATOR results and metadata

The two INTEGRATOR subsets share columns and hold disjoint analyses, so they are
combined by row-binding. The combined table is written to `output/` because it is
re-imported later (Figure 2) with MassHunter areas substituted in.

```{r chunk4-combine-import}
#| cache: false
sst_long   <- read_csv(file.path(d4, "Dataset4_MRMhub-INTEGRATOR_SST.csv"),   show_col_types = FALSE)
assay_long <- read_csv(file.path(d4, "Dataset4_MRMhub-INTEGRATOR_ASSAY.csv"), show_col_types = FALSE)
long_combined <- bind_rows(sst_long, assay_long)

long_combined_path <- file.path("output", "Dataset4_long_combined.csv")
write_csv(long_combined, long_combined_path)

# Analysis ids per subset (import strips the ".mzML" extension).
sst_analysis_ids   <- unique(str_remove(sst_long$raw_data_filename,   "[.]mzML$"))
assay_analysis_ids <- unique(str_remove(assay_long$raw_data_filename, "[.]mzML$"))

meta_path <- file.path(d4, "Dataset4_Metadata.xlsx")

mexp <- MRMhubExperiment(title = "Dataset 4 — Steroid Assay")
mexp <- import_data_mrmhub(mexp, path = long_combined_path, import_metadata = TRUE)
mexp <- import_metadata_msorganiser(
  mexp, path = meta_path, excl_unmatched_analyses = TRUE, ignore_warnings = TRUE)
```

## Normalization and calibration

Each feature is normalized to its stable-isotope-labelled internal standard, then
a quadratic, `1/x`-weighted external calibration curve is fitted per analyte from
the `Cal 0`–`Cal F` levels.

```{r chunk5-normalize-calibrate}
mexp <- normalize_by_istd(mexp)
mexp <- calc_calibration_results(
  mexp, fit_overwrite = TRUE, fit_model = "quadratic", fit_weighting = "1/x")
```

The fitted curves and their metrics are inspected before quantification.

```{r chunk6-calibration-curves}
#| fig-width: 9
#| fig-height: 7
#| fig-cap: "External calibration curves (quantifier transitions, quadratic / 1/x)."
#| fig.alt: "Calibration curves for the 15 steroid analytes."
plot_calibrationcurves(
  mexp, fit_overwrite = FALSE, fit_model = "quadratic", fit_weighting = "1/x",
  include_istd = FALSE, include_qualifier = FALSE, show_progress = FALSE)
```

```{r chunk7-calibration-metrics}
get_calibration_metrics(mexp) |>
  filter(is_quantifier) |>
  transmute(analyte = feature_id, fit_model, fit_weighting,
            r2 = round(r2, 5), lod = signif(lod, 3), loq = signif(loq, 3))
```

## Quantification

Concentrations are computed from the normalized intensities via the calibration
curves. Qualifier transitions are excluded; analytes whose curve failed to fit
are skipped rather than aborting the run.

```{r chunk8-quantify}
mexp <- quantify_by_calibration(
  mexp, fit_overwrite = FALSE, include_qualifier = FALSE,
  ignore_failed_calibration = TRUE, fit_model = "quadratic", fit_weighting = "1/x")

# Export + reload the concentration table (wide: analysis_id, qc_type, analytes).
conc_path <- file.path("output", "Dataset4_quant_results.csv")
save_dataset_csv(mexp, path = conc_path, variable = "conc", add_qctype = TRUE)
d_conc_raw <- read_csv(conc_path, show_col_types = FALSE)
```

## MassHunter re-integration

The MassHunter Quantitative Analysis export is imported as a second experiment
and reshaped to one tidy table — the single MassHunter source for all
comparisons below. The trailing ` (ISTD)` suffix on labelled internal standards
is stripped so feature identifiers match MRMhub.

```{r chunk9-masshunter}
mh_reint <- import_data_masshunter(
  MRMhubExperiment(title = "MassHunter re-integration"),
  path = file.path(d4, "Dataset4_MassHunter-V2.csv"),
  import_metadata = TRUE, silent = TRUE)
mh_data <- get_analyticaldata(mh_reint, annotated = FALSE) |>
  transmute(analysis_id,
            feature_id = str_remove(feature_id, "[ ]*\\(ISTD\\)$"),
            rt_mh   = feature_rt,
            area_mh = feature_area,
            conc_mh = feature_conc_final)
```

## Below-Cal-A exclusion for study samples

Study samples (`SPL` = the SKML EQA samples) whose **quantifier peak area falls
below the corresponding `Cal A` area** are below the lowest non-zero calibrator
and treated as not reliably quantifiable. The rule is applied per analyte and
**independently per method** — MRMhub SPL area vs MRMhub Cal A area, MassHunter
SPL area vs MassHunter Cal A area — yielding the flags `excl_mrm` / `excl_mh` that
downstream comparisons apply to their SPL rows. Internal standards are never
filtered.

```{r chunk10-spl-exclusion}
CAL_A_ID <- "Cal A_P1"

# MRMhub quantifier areas (quantifiers only, excluding the ISTDs).
mrm_q_area <- get_analyticaldata(mexp, annotated = TRUE) |>
  filter(is_quantifier, !is_istd) |>
  transmute(analysis_id, analyte_id = feature_id, qc_type, area_mrm = feature_area)

cala_mrm <- mrm_q_area |>
  filter(analysis_id == CAL_A_ID) |>
  select(analyte_id, cala_area_mrm = area_mrm)
cala_mh <- mh_data |>
  filter(analysis_id == CAL_A_ID) |>
  transmute(analyte_id = feature_id, cala_area_mh = area_mh)

# One row per SPL (analysis, analyte); flag below-Cal-A per method (NA-safe).
spl_area_excl <- mrm_q_area |>
  filter(qc_type == "SPL") |>
  left_join(select(mh_data, analysis_id, analyte_id = feature_id, area_mh),
            by = c("analysis_id", "analyte_id")) |>
  left_join(cala_mrm, by = "analyte_id") |>
  left_join(cala_mh,  by = "analyte_id") |>
  mutate(excl_mrm = !is.na(area_mrm) & !is.na(cala_area_mrm) & area_mrm < cala_area_mrm,
         excl_mh  = !is.na(area_mh)  & !is.na(cala_area_mh)  & area_mh  < cala_area_mh) |>
  select(analysis_id, analyte_id, area_mrm, area_mh,
         cala_area_mrm, cala_area_mh, excl_mrm, excl_mh)
```

## Low / high QC

Quantified concentrations of the low and high QC samples, one row per analyte.

```{r chunk11-qc-highlow}
d_conc_raw |>
  filter(qc_type == "QC") |>
  pivot_longer(!c(analysis_id, qc_type), names_to = "analyte", values_to = "conc") |>
  mutate(qc = if_else(str_detect(analysis_id, "High"), "QC_High", "QC_Low")) |>
  select(analyte, qc, conc) |>
  pivot_wider(names_from = qc, values_from = conc) |>
  mutate(across(where(is.numeric), ~ signif(.x, 3)))
```

# MRMhub vs MassHunter — Peak Integration

Figure 1 compares MRMhub-INTEGRATOR peak areas with the MassHunter
re-integration of the same raw files. MRMhub quantifier and ISTD peaks are joined
to MassHunter on `analysis_id` + `feature_id`, dropping near-noise peaks
(area < 100), empty CAL rows, and below-Cal-A SPL points.

```{r chunk12-house-style}
# Manuscript house theme + palettes (shared by both figures).
theme_set(theme_bw(base_size = 8, base_family = "sans") +
            theme(panel.grid.minor = element_blank(),
                  legend.position  = "bottom",
                  plot.title       = element_text(size = 8, face = "plain"),
                  legend.key.size  = unit(3.2, "mm"),
                  legend.margin    = margin(1, 1, 1, 1)))

col_method <- c(MRMhub = "#2C7FB8", MassHunter = "#D95F0E")
smp_cols   <- c(CAL = "#0072B2", HQC = "#D55E00", LQC = "#E69F00",
                SPL = "grey30", SST = "#009E73", SBLK = "#CC79A7", PBLK = "#999999")

# Save a composite as vector PDF + 300-dpi PNG at a given size (mm) for the manuscript.
save_fig <- function(plot, stem, height, width = 180) {
  ggsave(file.path("output", paste0(stem, ".pdf")), plot,
         device = cairo_pdf, width = width, height = height, units = "mm")
  ggsave(file.path("output", paste0(stem, ".png")), plot,
         device = ragg::agg_png, width = width, height = height, units = "mm", dpi = 300)
  invisible(plot)
}
```

```{r chunk13-fig1-data}
mrm_peaks_q <- get_analyticaldata(mexp, annotated = TRUE) |>
  filter(is_quantifier | is_istd) |>
  transmute(analysis_id, feature_id, qc_type, sample_id,
            feature_type = factor(if_else(is_istd, "ISTD", "Quantifier"),
                                  levels = c("Quantifier", "ISTD")),
            rt_mrm = feature_rt, area_mrm = feature_area)

d_peakcmp_reint <- mrm_peaks_q |>
  inner_join(mh_data, by = c("analysis_id", "feature_id")) |>
  filter(!is.na(rt_mrm), !is.na(rt_mh), !is.na(area_mrm), !is.na(area_mh)) |>
  filter(area_mrm >= 100, area_mh >= 100) |>
  filter(!(qc_type == "CAL" & is.na(sample_id) & feature_type != "ISTD")) |>
  left_join(select(spl_area_excl, analysis_id, analyte_id, excl_mrm, excl_mh),
            by = c("analysis_id", "feature_id" = "analyte_id")) |>
  filter(!(qc_type == "SPL" &
             (coalesce(excl_mrm, FALSE) | coalesce(excl_mh, FALSE)))) |>
  select(-excl_mrm, -excl_mh) |>
  mutate(area_diff_pct = (area_mh - area_mrm) / area_mrm * 100,
         rt_diff = rt_mh - rt_mrm,
         subset = case_when(
           analysis_id %in% sst_analysis_ids   ~ "SST",
           analysis_id %in% assay_analysis_ids ~ "ASSAY",
           TRUE                                ~ NA_character_))
```

## Peak-area agreement

Log–log peak areas against the 1:1 line (Pearson *r* on log10 values). CAL/QC/SST
samples span four orders of magnitude; agreement is essentially exact.

```{r chunk14-fig1-panelA}
#| fig-width: 4.6
#| fig-height: 4.2
smp_lv    <- c("CAL", "HQC", "LQC", "SST")
smp_cols4 <- c(CAL = "#0072B2", HQC = "#D55E00", LQC = "#E69F00", SST = "#009E73")
ft_shapes <- c(Quantifier = 16, ISTD = 17)
scale_smp <- scale_colour_manual(values = smp_cols4, limits = smp_lv, drop = FALSE)
scale_ft  <- scale_shape_manual(values = ft_shapes)

d_cmp   <- d_peakcmp_reint |>
  filter(qc_type %in% smp_lv) |>
  mutate(qc_type = factor(qc_type, levels = smp_lv))
r_log   <- cor(log10(d_cmp$area_mrm), log10(d_cmp$area_mh))
ax_lims <- range(c(d_cmp$area_mrm, d_cmp$area_mh))
p_corr <- d_cmp |>
  ggplot(aes(area_mrm, area_mh, colour = qc_type, shape = feature_type)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
  geom_point(size = 1.1, alpha = 0.7) +
  annotate("text", x = ax_lims[1], y = ax_lims[2], hjust = 0, vjust = 1, size = 2.4,
           label = sprintf("Pearson r = %.4f", r_log)) +
  scale_x_log10(limits = ax_lims) + scale_y_log10(limits = ax_lims) +
  scale_smp + scale_ft +
  guides(colour = guide_legend(order = 1, nrow = 2),
         shape  = guide_legend(order = 2, nrow = 2)) +
  labs(x = "MRMhub peak area", y = "MassHunter peak area",
       colour = "Sample type", shape = "Feature", title = "Peak-area agreement")
p_corr
```

## Precision (%CV) equivalence

Per-feature %CV over the ten replicate SST injections, MRMhub vs MassHunter. The
two methods deliver equivalent precision (points on the 1:1 line).

```{r chunk15-fig1-panelB}
#| fig-width: 4.6
#| fig-height: 4.2
sst_cv <- d_peakcmp_reint |>
  filter(subset == "SST") |>
  group_by(feature_type, feature_id) |>
  summarise(cv_mrmhub     = sd(area_mrm) / mean(area_mrm) * 100,
            cv_masshunter = sd(area_mh)  / mean(area_mh)  * 100,
            .groups = "drop")
cv_lims <- range(c(sst_cv$cv_mrmhub, sst_cv$cv_masshunter)) + c(-0.3, 0.3)
p_cv <- sst_cv |>
  ggplot(aes(cv_mrmhub, cv_masshunter, shape = feature_type)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey60") +
  geom_point(size = 1.6, alpha = 0.85, colour = "#009E73") +
  coord_cartesian(xlim = cv_lims, ylim = cv_lims) + scale_ft +
  guides(shape = "none") +
  labs(x = "MRMhub %CV (10 SST reps)", y = "MassHunter %CV",
       title = "Precision (%CV) equivalence")
p_cv
```

## Bland–Altman

Percent difference (MassHunter − MRMhub) against the mean area, excluding SST and
near-noise peaks (mean area < 10³). The three analytes highlighted here have the
largest real-sample differences and their chromatograms are shown below.

```{r chunk16-fig1-panelC}
#| fig-width: 7
#| fig-height: 3.4
highlight_analytes <- c("21-deoxycortisol", "11-deoxycortisol", "Aldosterone")
hl_cols <- setNames(c("#D81B60", "#1E88E5", "#8E44AD"), highlight_analytes)

d_ba <- d_peakcmp_reint |>
  filter(qc_type != "SST") |>
  mutate(mean_area = (area_mrm + area_mh) / 2,
         pct_diff  = (area_mh - area_mrm) / mean_area * 100,
         hl = factor(if_else(feature_id %in% highlight_analytes, feature_id, NA_character_),
                     levels = highlight_analytes)) |>
  filter(mean_area >= 1e3)
ba_bias <- mean(d_ba$pct_diff)
ba_loa  <- ba_bias + c(-1.96, 1.96) * sd(d_ba$pct_diff)
ba_xr   <- max(d_ba$mean_area)
p_ba <- d_ba |>
  ggplot(aes(mean_area, pct_diff)) +
  geom_hline(yintercept = 0, colour = "grey70") +
  geom_hline(yintercept = ba_bias, colour = "grey20") +
  geom_hline(yintercept = ba_loa, linetype = "dashed", colour = "grey50") +
  geom_point(data = ~ filter(.x, is.na(hl)), colour = "grey80", size = 0.9, alpha = 0.6) +
  geom_point(data = ~ filter(.x, !is.na(hl)), aes(colour = hl), size = 1.7, alpha = 0.9) +
  annotate("text", x = ba_xr, y = ba_bias,  hjust = 1, vjust = -0.6, size = 2.1,
           colour = "grey20", label = sprintf("mean bias %.1f%%", ba_bias)) +
  annotate("text", x = ba_xr, y = ba_loa[2], hjust = 1, vjust = 1.4, size = 2,
           colour = "grey45", label = "+95% LoA") +
  annotate("text", x = ba_xr, y = ba_loa[1], hjust = 1, vjust = -0.7, size = 2,
           colour = "grey45", label = "-95% LoA") +
  scale_colour_manual(values = hl_cols, name = "Highlighted analyte", na.translate = FALSE) +
  scale_x_log10(limits = c(1e3, NA)) + coord_cartesian(ylim = c(-20, 20)) +
  labs(x = "Mean peak area of the two methods (>= 10^3)",
       y = "Percent difference (MH - MRMhub) %",
       title = sprintf("Bland-Altman (excl. SST, area >= 10^3): mean bias %.1f%%, 95%% LoA [%.0f, %.0f]%%",
                       ba_bias, ba_loa[1], ba_loa[2]))
p_ba
```

## Example chromatograms

The largest real-sample area differences occur where a peak sits on an elevated
baseline or beside a co-eluting neighbour, so the integration boundaries — and
hence the areas — legitimately differ between the two integrators. The candidate
table ranks these cases; the panel crops the three chosen examples (one per
highlighted analyte) from the MRMhub `by_transition` chromatograms.

```{r chunk17-fig1-chrom-candidates}
chrom_candidates <- d_peakcmp_reint |>
  filter(feature_type == "Quantifier",
         qc_type %in% c("CAL", "HQC", "LQC", "SPL", "SST"),
         area_mrm >= 20000) |>
  arrange(desc(abs(area_diff_pct))) |>
  transmute(feature_id, analysis_id, qc_type,
            area_mrm = round(area_mrm), area_mh = round(area_mh),
            area_diff_pct = round(area_diff_pct, 1)) |>
  head(10)
chrom_candidates
```

```{r chunk18-fig1-chrom-panel}
#| fig-width: 8
#| fig-height: 3
# Panel-D chromatograms are the pre-generated crops (see the INTEGRATOR section);
# subtitles carry the data-derived area difference for each example.
chrom_examples <- tibble::tribble(
  ~name,                    ~feature_id,        ~analysis_id,
  "fig1d_21-deoxycortisol", "21-deoxycortisol", "SKML2025_6B_P1",
  "fig1d_11-deoxycortisol", "11-deoxycortisol", "SKML2026_2B_P1",
  "fig1d_Aldosterone",      "Aldosterone",      "SKML2026_2A_P1")

chrom_meta <- chrom_examples |>
  left_join(distinct(d_peakcmp_reint, feature_id, analysis_id, area_diff_pct),
            by = c("feature_id", "analysis_id")) |>
  mutate(subtitle = sprintf("%s  (dArea %+.1f%%)",
                            str_remove(analysis_id, "_P1$"), area_diff_pct))
chrom_plots <- Map(chrom_gg, chrom_meta$name, chrom_meta$feature_id, chrom_meta$subtitle)
p_chrom <- wrap_elements(full = wrap_plots(chrom_plots, nrow = 1))
p_chrom
```

## Assembled figure for the manuscript

The panels above are composed into the publication figure and saved as a vector
PDF and a 300-dpi PNG (Nature 180 mm width).

```{r chunk19-fig1-assemble}
#| fig-width: 7.1
#| fig-height: 9.06
row_ab <- (p_corr + p_cv) + plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
fig1 <- row_ab / p_ba / p_chrom +
  plot_layout(heights = c(1, 0.72, 0.58)) +
  plot_annotation(tag_levels = "A")

save_fig(fig1, "fig1_integration_correspondence", height = 230)
fig1
```

# MRMhub vs MassHunter — Quantification

Figure 2 compares the two pipelines at the quantification stage. To isolate the
post-processing (calibration + quantification) from the peak integration, we also
feed **MassHunter's own peak areas through the MRMhub pipeline** — identical
areas, identical calibration model — so any concentration difference is due to
post-processing alone.

```{r chunk20-fig2-mhpipe}
mh_area <- mh_data |> transmute(analysis_id, feature_name = feature_id, area_mh)
long_mh <- read_csv(long_combined_path, show_col_types = FALSE) |>
  mutate(.aid = str_remove(raw_data_filename, "[.]mzML$")) |>
  inner_join(mh_area, by = c(".aid" = "analysis_id", "feature_name")) |>
  mutate(area = area_mh) |>
  select(-.aid, -area_mh)

tmp_long <- tempfile(fileext = ".csv")
write_csv(long_mh, tmp_long)

mexp_mh <- MRMhubExperiment(title = "MassHunter areas via MRMhub pipeline")
mexp_mh <- import_data_mrmhub(mexp_mh, path = tmp_long, import_metadata = TRUE)
mexp_mh <- import_metadata_msorganiser(
  mexp_mh, path = meta_path, excl_unmatched_analyses = TRUE, ignore_warnings = TRUE)
mexp_mh <- normalize_by_istd(mexp_mh)
mexp_mh <- calc_calibration_results(mexp_mh, fit_overwrite = TRUE,
                                    fit_model = "quadratic", fit_weighting = "1/x")
mexp_mh <- quantify_by_calibration(mexp_mh, fit_overwrite = FALSE,
                                   include_qualifier = FALSE,
                                   ignore_failed_calibration = TRUE,
                                   fit_model = "quadratic", fit_weighting = "1/x")
```

## Calibration fit comparison

Calibration R² per analyte: MassHunter's own fit vs the MRMhub full-pipeline fit.

```{r chunk21-fig2-panelA}
#| fig-width: 6
#| fig-height: 4.5
mh_own_r2 <- read_csv(file.path(d4, "Dataset4_MassHunter-CalibrationFits.csv"),
                      show_col_types = FALSE) |>
  transmute(analyte_id, r2_mh_own = r2)
r2_cmp <- get_calibration_metrics(mexp_mh) |>
  filter(is_quantifier) |>
  transmute(analyte_id = feature_id, r2_mrmpipe_mharea = r2) |>
  left_join(mh_own_r2, by = "analyte_id") |>
  left_join(get_calibration_metrics(mexp) |> filter(is_quantifier) |>
              transmute(analyte_id = feature_id, r2_mrmpipe_mrmarea = r2),
            by = "analyte_id")

chain_lv  <- c("MassHunter", "MRMhub")
chain_col <- c(MassHunter = "#D95F0E", MRMhub = "#2C7FB8")
ord_r2 <- r2_cmp |> arrange(r2_mrmpipe_mrmarea) |> pull(analyte_id)

d_r2_long <- r2_cmp |>
  transmute(analyte_id, MassHunter = r2_mh_own, MRMhub = r2_mrmpipe_mrmarea) |>
  pivot_longer(-analyte_id, names_to = "chain", values_to = "r2") |>
  mutate(analyte_id = factor(analyte_id, levels = ord_r2),
         chain = factor(chain, levels = chain_lv))

# Truncate the x-axis just below the dense 0.996–1.000 cluster; flag off-scale rows.
R2_FLOOR <- 0.9955
r2_off <- d_r2_long |> filter(r2 < R2_FLOOR) |> distinct(analyte_id) |>
  mutate(r2 = R2_FLOOR)
off_txt <- d_r2_long |> filter(r2 < R2_FLOOR) |>
  group_by(analyte_id) |> summarise(v = min(r2), .groups = "drop") |>
  arrange(v) |> mutate(t = sprintf("%s %.3f", analyte_id, v))
p_r2 <- ggplot(d_r2_long, aes(r2, analyte_id)) +
  geom_line(aes(group = analyte_id), colour = "grey75", linewidth = 0.6) +
  geom_point(aes(colour = chain), size = 1.9, alpha = 0.9) +
  geom_text(data = r2_off, aes(label = "<"), hjust = 0, nudge_x = 0.00004,
            size = 2.3, colour = "grey45") +
  scale_colour_manual(values = chain_col, name = NULL) +
  scale_x_continuous(breaks = c(0.996, 0.997, 0.998, 0.999, 1.000)) +
  coord_cartesian(xlim = c(R2_FLOOR, 1.0004)) +
  labs(x = expression(Calibration~R^2), y = NULL, title = "Calibration curve fit comparison",
       subtitle = paste0("x-axis truncated at ", R2_FLOOR,
                         "; off scale (<): ", paste(off_txt$t, collapse = ", "))) +
  theme(plot.subtitle = element_text(size = 6.5, colour = "grey30"),
        plot.margin = margin(4, 10, 4, 4),
        legend.position = "inside", legend.position.inside = c(0.015, 0.98),
        legend.justification.inside = c(0, 1),
        legend.background = element_rect(fill = alpha("white", 0.7), colour = NA),
        legend.key.size = unit(3, "mm"))
p_r2
```

## Concentration equivalence on identical areas

Feeding MassHunter areas through the MRMhub pipeline reproduces MassHunter's
end-to-end concentrations almost exactly — the post-processing steps are
equivalent.

```{r chunk22-fig2-panelB}
#| fig-width: 4.6
#| fig-height: 4.4
tmp_conc <- tempfile(fileext = ".csv")
save_dataset_csv(mexp_mh, path = tmp_conc, variable = "conc", add_qctype = TRUE)
d_qonly <- read_csv(tmp_conc, show_col_types = FALSE) |>
  pivot_longer(!c(analysis_id, qc_type), names_to = "analyte_id",
               values_to = "conc_mrmpipe") |>
  left_join(mh_data |> select(analysis_id, analyte_id = feature_id, conc_mh),
            by = c("analysis_id", "analyte_id")) |>
  filter(!is.na(conc_mrmpipe), !is.na(conc_mh)) |>
  filter(!qc_type %in% c("SBLK", "PBLK")) |>
  mutate(qc_type = factor(qc_type, levels = names(smp_cols)))
posq  <- d_qonly |> filter(conc_mrmpipe > 0, conc_mh > 0)
q_r   <- cor(log10(posq$conc_mrmpipe), log10(posq$conc_mh))
q_med <- median(abs((d_qonly$conc_mrmpipe - d_qonly$conc_mh) / d_qonly$conc_mh * 100))
axq   <- range(c(posq$conc_mrmpipe, posq$conc_mh))
p_qonly <- posq |>
  ggplot(aes(conc_mh, conc_mrmpipe, colour = qc_type)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
  geom_point(size = 1.1, alpha = 0.6) +
  annotate("text", x = axq[1], y = axq[2], hjust = 0, vjust = 1, size = 2.4,
           label = sprintf("Pearson r = %.5f\nmedian |d| = %.3f%%", q_r, q_med)) +
  scale_x_log10(limits = axq) + scale_y_log10(limits = axq) +
  scale_colour_manual(values = smp_cols, drop = TRUE, name = "Sample type") +
  coord_equal() +
  labs(x = "MassHunter End-to-End", y = "MRMhub-QUANT on MH areas",
       title = "Concentration equivalence (same areas)") +
  theme(legend.position = "inside", legend.position.inside = c(0.98, 0.02),
        legend.justification.inside = c(1, 0),
        legend.background = element_rect(fill = alpha("white", 0.7), colour = NA),
        legend.key.size = unit(2.8, "mm"),
        legend.title = element_text(size = 6.5), legend.text = element_text(size = 6))
p_qonly
```

## Full end-to-end concentration agreement

Independent end-to-end concentrations (each method's own integration and
calibration) for the LQC, HQC and SPL samples.

```{r chunk23-fig2-panelC}
#| fig-width: 4.6
#| fig-height: 4.4
d_full <- d_conc_raw |>
  filter(qc_type %in% c("LQC", "HQC", "SPL")) |>
  pivot_longer(!c("analysis_id", "qc_type"), names_to = "analyte_id", values_to = "conc") |>
  left_join(mh_data |> select(analysis_id, analyte_id = feature_id, conc_mh),
            by = c("analysis_id", "analyte_id")) |>
  left_join(select(spl_area_excl, analysis_id, analyte_id, excl_mrm, excl_mh),
            by = c("analysis_id", "analyte_id")) |>
  mutate(conc    = if_else(coalesce(excl_mrm, FALSE), NA_real_, conc),
         conc_mh = if_else(coalesce(excl_mh,  FALSE), NA_real_, conc_mh)) |>
  select(-excl_mrm, -excl_mh) |>
  filter(!is.na(conc), !is.na(conc_mh), conc > 0, conc_mh > 0) |>
  mutate(qc_type = factor(qc_type, levels = names(smp_cols)))
f_r   <- cor(log10(d_full$conc), log10(d_full$conc_mh))
f_med <- median(abs((d_full$conc - d_full$conc_mh) / d_full$conc_mh * 100))
axf   <- range(c(d_full$conc, d_full$conc_mh))
p_full <- d_full |>
  ggplot(aes(conc_mh, conc, colour = qc_type)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
  geom_point(size = 1.3, alpha = 0.7) +
  annotate("text", x = axf[1], y = axf[2], hjust = 0, vjust = 1, size = 2.4,
           label = sprintf("Pearson r = %.4f\nmedian |d| = %.1f%%", f_r, f_med)) +
  scale_x_log10(limits = axf) + scale_y_log10(limits = axf) +
  scale_colour_manual(values = smp_cols, drop = TRUE, name = "Sample type") +
  coord_equal() +
  labs(x = "MassHunter (end-to-end)", y = "MRMhub (end-to-end)",
       title = "Concentration agreement (full pipeline)") +
  theme(legend.position = "none")
p_full
```

## EQA accuracy against SKML reference values

Per-analyte percent bias against the SKML EQA reference concentrations, one point
per EQA sample per method (dashed lines: ±15 %). The reference concentrations are
read from the metadata workbook's `QC Concentrations` sheet (imported above with
the rest of the metadata), keyed by sample and analyte. DHEA and DHEAS are excluded
(large biases that would compress the other panels).

```{r chunk24-fig2-panelD}
#| fig-width: 7
#| fig-height: 4.5
# EQA target concentrations come from the metadata's `QC Concentrations` sheet
# (imported above), keyed by `sample_id` + `analyte_id`. They surface via the
# `annot_qcconcentrations` slot; DHEAS is already stored in nmol/L there.
eqa_target <- mexp@annot_qcconcentrations |>
  filter(str_starts(sample_id, "SKML")) |>
  transmute(sample_id, analyte_id, target = concentration)
# analysis_id -> sample_id map (groups the `…b` replicates onto one EQA sample).
sid_map <- get_analyticaldata(mexp, annotated = TRUE) |>
  distinct(analysis_id, sample_id)
d_eqa <- d_conc_raw |>
  filter(qc_type == "SPL") |>
  pivot_longer(!c("analysis_id", "qc_type"), names_to = "analyte_id", values_to = "conc_mrm") |>
  left_join(mh_data |> select(analysis_id, analyte_id = feature_id, conc_mh),
            by = c("analysis_id", "analyte_id")) |>
  left_join(select(spl_area_excl, analysis_id, analyte_id, excl_mrm, excl_mh),
            by = c("analysis_id", "analyte_id")) |>
  mutate(conc_mrm = if_else(coalesce(excl_mrm, FALSE), NA_real_, conc_mrm),
         conc_mh  = if_else(coalesce(excl_mh,  FALSE), NA_real_, conc_mh)) |>
  left_join(sid_map, by = "analysis_id") |>
  inner_join(eqa_target, by = c("sample_id", "analyte_id")) |>
  # Compact axis label, e.g. SKML2026_1A -> 2026.1A.
  mutate(sample_eqa = str_replace(str_remove(sample_id, "^SKML"), "^([0-9]{4})_", "\\1.")) |>
  transmute(analyte_id, sample_eqa, analysis_id, target, conc_mrm, conc_mh,
            bias_mrmhub     = (conc_mrm - target) / target * 100,
            bias_masshunter = (conc_mh  - target) / target * 100) |>
  filter(!(is.na(conc_mrm) & is.na(conc_mh)))
p_eqa <- d_eqa |>
  filter(!analyte_id %in% c("DHEA", "DHEAS")) |>
  pivot_longer(c(bias_mrmhub, bias_masshunter), names_to = "method", values_to = "bias") |>
  filter(!is.na(bias)) |>
  mutate(method = recode(method, bias_mrmhub = "MRMhub", bias_masshunter = "MassHunter"),
         method = factor(method, levels = c("MRMhub", "MassHunter"))) |>
  ggplot(aes(sample_eqa, bias, colour = method, shape = method)) +
  geom_hline(yintercept = 0, colour = "grey40") +
  geom_hline(yintercept = c(-15, 15), linetype = "dashed", colour = "grey75") +
  geom_point(size = 1.5, alpha = 0.85, position = position_dodge(width = 0.4)) +
  scale_colour_manual(values = col_method, name = "Method") +
  scale_shape_manual(values = c(MRMhub = 16, MassHunter = 17), name = "Method") +
  facet_wrap(vars(analyte_id), scales = "free_y", ncol = 4) +
  labs(x = NULL, y = "Bias vs EQA reference (%)",
       title = "EQA accuracy - bias per analyte (excl. DHEA, DHEAS)") +
  theme(panel.grid.minor = element_blank(),
        strip.text = element_text(size = 6),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 5),
        axis.text.y = element_text(size = 5.2))
p_eqa
```

## Assembled figure for the manuscript

```{r chunk25-fig2-assemble}
#| fig-width: 7.1
#| fig-height: 8.66
fig2 <- p_r2 / (p_qonly + p_full) / p_eqa +
  plot_layout(heights = c(0.68, 0.95, 1.2)) +
  plot_annotation(tag_levels = "A")

save_fig(fig2, "fig2_quantification_comparison", height = 220)
fig2
```

# References
