---
title: "Dataset 4 — Quantification & QC Report"
subtitle: "Steroid Assay"
author:
  - name: "Bo Burla"
    affiliation: "National University of Singapore"
  - name: "Guo Shou Teo"
  - name: "Hyungwon Choi"
date: "`r Sys.Date()`"
format:
  html:
    code-tools: true
    self-contained: true
    code-overflow: wrap
    toc: true
    number-sections: true
    df-print: paged
    lightbox: true
    fig-align: center
    fig-format: retina
    fig-dpi: 200
bibliography: references.bib
---

# Overview

This is an MRMhub postprocessing report for Dataset 4, a **fully quantitative 15-analyte serum steroid panel** measured with external calibration. The data are from an LC-MRM-MS analysis using the Steroid Panel LC-MS kit (IBL International, Germany), following the Instructions for Use except that LC-MS was performed on an Agilent 6495A triple quadrupole instrument. Only the Gradient 1 analysis was used, covering 15 analytes with their internal standards (66 MRM transitions monitored, including qualifiers, from which 64 features were integrated).

The dataset comprises system-suitability samples (SST, containing pure analyte and internal standards), blanks, calibrants (Cal A – Cal F), low and high quality controls (LQC, HQC) with assigned target concentrations, and, in place of study samples, 5 external quality assessment (EQA) samples (SKML [@jansen2014]).

::: callout-important
## Data availability

The datasets and the full code used to generate this resource are deposited in the Zenodo record [10.5281/zenodo.15370293](https://doi.org/10.5281/zenodo.15370293) (**MRMhub-workflows**).
:::

::: callout-note
## mrmhub version

Rendered with **mrmhub 0.9.9** (QUANT module of [MRMhub](https://github.com/SLINGhub/MRMhub), `main` branch): `pak::pak("SLINGhub/MRMhub")`.
:::

## Setup

Load the package and call `mrmhub_enable_cli_color()` to render mrmhub's console feedback in colour in the HTML output.

```{r chunk2-loadpkg}
library(mrmhub)
mrmhub_enable_cli_color()
```

# Import

Data and corresponding metadata are loaded into a fresh `MRMhubExperiment`. `import_data_mrmhub()` reads the INTEGRATOR long-format peak areas for the assay batch (blanks, calibrators, QCs and EQA samples). `import_metadata_msorganiser()` then attaches mrmhub's metadata workbook template, which defines the analytes, their internal standards, the sample annotations, and the calibration/target concentrations.

```{r chunk3-import}
mexp <- MRMhubExperiment(title = "Dataset 4 — Steroid Assay")
mexp <- import_data_mrmhub(
  mexp,
  path = "./data/dataset-4/Dataset4_MRMhub-INTEGRATOR_ASSAY.csv",
  import_metadata = TRUE
)
mexp <- import_metadata_msorganiser(
  mexp,
  path = "./data/dataset-4/Dataset4_Metadata.xlsx",
  excl_unmatched_analyses = TRUE, ignore_warnings = TRUE
)
```

# Calibration

`normalize_by_istd()` divides each analyte's peak area by that of its authentic stable-isotope-labelled internal standard. `calc_calibration_results()` then fits one external calibration curve per analyte from the `Cal A`–`Cal F` levels. A **quadratic** model with **`1/x` weighting** was used as the regression model for all analytes.

```{r chunk4-normalize-calibrate}
mexp <- normalize_by_istd(mexp)
mexp <- calc_calibration_results(
  mexp,
  fit_overwrite = TRUE,
  fit_model = "quadratic",
  fit_weighting = "1/x"
)
```

```{r}
#| label: fig-d4rep-calibration-curves
#| fig-cap: "**External calibration curves.** Quadratic, 1/x-weighted curves for the 15 steroid analytes (quantifier transitions)."
#| fig-width: 9
#| fig-height: 7
#| out-width: "100%"
#| fig.alt: "Calibration curves for the 15 steroid analytes."
# Curves are solid between the lowest and highest calibrator, and dotted where
# extrapolated beyond that calibrator range (automatic; not a range restriction).
plot_calibrationcurves(
  mexp,
  fit_overwrite = FALSE,
  fit_model = "quadratic",
  fit_weighting = "1/x",
  include_istd = FALSE,
  include_qualifier = FALSE,
  rows_page = 4,
  cols_page = 4,
  show_progress = FALSE
)
```

The metrics below summarize each quantifier's fit: **R²** (linearity/goodness of fit) and the estimated **LOD** and **LOQ** (limits of detection and quantification, in the calibrated concentration unit).

```{r}
#| label: tbl-d4rep-calibration-metrics
#| tbl-cap: "**Calibration metrics per analyte.**"
get_calibration_metrics(mexp, include_qualifier = FALSE, summary_table = TRUE)
```

# Quantification

`quantify_by_calibration()` inverts each fitted curve to convert the normalized intensities into absolute concentrations. Only quantifier transitions are used (`include_qualifier = FALSE`), and `ignore_failed_calibration = TRUE` skips any analyte whose curve failed to fit rather than aborting the whole run, so a single problematic analyte does not block the report. `fit_model` and `fit_weighting` can also be set per analyte in the feature metadata table.

```{r chunk7-quantify}
mexp <- quantify_by_calibration(
  mexp,
  fit_overwrite = FALSE,
  include_qualifier = FALSE,
  ignore_failed_calibration = TRUE,
  fit_model = "quadratic",
  fit_weighting = "1/x"
)
```

# QC bias & variability

The final check compares the measured QC concentrations against their known targets. `get_qc_bias_variability()` reports, per analyte and QC level, the mean measured concentration, the **bias** (percent deviation from target — *accuracy*) and the intra-batch **%CV** (*precision*), plus a flag for values falling outside the calibrated range. Restricting to `LQC`/`HQC` keeps the table to the samples that have assigned target concentrations.

```{r}
#| label: tbl-d4rep-qc-bias-variability
#| tbl-cap: "**QC bias and variability** for the LQC and HQC samples."
get_qc_bias_variability(mexp, qc_types = c("LQC", "HQC"), summary_table = TRUE)
```

# Export

The results are written in three complementary formats, each for a different use.

## Flat concentration table (CSV)

`save_dataset_csv()` writes a single flat table of concentrations, one row per sample — the simplest format to open in Excel or read into any tool. `add_qctype` keeps the QC-type column so calibrators, QCs and samples stay distinguishable.

```{r chunk9-export-csv}
save_dataset_csv(
  mexp,
  path = file.path("output", "Dataset4_report_conc.csv"),
  variable = "conc",
  add_qctype = TRUE
)
```

## Multi-sheet report (XLSX)

`save_report_xlsx()` writes a multi-sheet workbook bundling the concentrations *and* the supporting context (calibration metrics, feature QC metrics, raw/normalized intensities, sample and feature metadata) in one file for sharing and archiving.

```{r chunk10-export-xlsx}
save_report_xlsx(
  mexp,
  path = file.path("output", "Dataset4_report.xlsx"),
  filtered_variable = "conc",
  overwrite = TRUE
)
```

## Complete dataset object (RDS)

`save_dataset_rds()` writes the entire `MRMhubExperiment` — imported data, metadata, calibration fits and quantification results — to a single `.rds`, so that everything needed to reproduce, extend or audit the analysis is contained in one file. An `rlang::hash()` fingerprint is embedded, allowing the file's integrity to be verified when it is read back. The file is a plain R object and can therefore be opened in any R session with `readRDS()`, even without mrmhub installed; with mrmhub loaded, the object's methods are available and the analysis can be continued from where this report ends.

```{r chunk11-export-rds}
save_dataset_rds(mexp, "output/Dataset4_report_mexp.rds")
```

Reload it in a fresh R session with `read_dataset_rds()`, which verifies the embedded integrity hash on read (load `mrmhub` first so the object's methods are available):

```{r chunk12-rds-import}
#| eval: false
library(mrmhub)
mexp <- read_dataset_rds("output/Dataset4_report_mexp.rds")
```

# References