---
title: "MRMhub Data Processing Workflow for Dataset 4"
subtitle: "Supplementary Code 3"
author:
  - name: "Bo Burla"
    affiliation: "National University of Singapore"
  - name: "Guo Shou Teo"
  - name: "Hyungwon Choi"
date: "`r Sys.Date()`"
format:
  html:
    code-tools: true
    code-fold: show
    self-contained: true
    code-overflow: wrap
    toc: true
    number-sections: true
    df-print: paged
    lightbox: true
    fig-align: center
    fig-format: retina
    fig-dpi: 200
  pdf:
    pdf-engine: lualatex
    mainfont: "Latin Modern Sans"
    fig-format: pdf
    toc: true
    fig-dpi: 300
    number-sections: true
    keep-tex: false
    include-in-header: caption-font.tex
bibliography: references.bib
---

```{r chunk1-init, include = FALSE, echo = FALSE}
#| cache: false
# Rendering options for this document. Caching is off: the dataset is small
# (15 analytes, ~31 analyses) so a full render is fast, and several chunks write
# intermediate files, where stale cache state would be a liability.
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE, results = TRUE, message = TRUE, warning = TRUE,
  comment = "#>", out.width = "100%"
)
set.seed(1041)
options(dplyr.print_max = 15)

library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(ggplot2)
library(patchwork)

# Data root for this notebook (flat layout, mirroring Dataset 1/3).
d4 <- "./data/dataset-4"
dir.create("output", showWarnings = FALSE, recursive = TRUE)
```

\newpage

# Overview

Dataset 4 is a **fully quantitative steroid panel** used to validate MRMhub
against a vendor pipeline (Agilent MassHunter). Fifteen steroids were measured in
human serum with the IBL International *Steroid Panel LC-MS* kit
(cat. 30191875) [@ibl_steroidpanel] according to the instructions for use, with
modifications. The
kit acquires its analytes over two chromatographic panels; **only Panel 1 (15
analytes) is used here**, which is why every sample identifier carries the `_P1`
suffix (e.g. `Cal A_P1`, `QC_High1_P1`).

Unlike Datasets 1 and 3 (relative lipidomics), this is an externally calibrated
absolute assay. The analytical design comprises an **external calibration curve**
(`Cal 0`–`Cal F`), **low and high quality-control samples** (`QC_Low1`,
`QC_High1`), and **external quality-assessment (EQA) samples** — SKML ring-trial
materials with reference-method-assigned target concentrations [@jansen2014].

The workflow compares MRMhub and MassHunter at both pipeline stages:

- **INTEGRATOR** — peak areas from MRMhub-INTEGRATOR vs a MassHunter
  re-integration of the same raw files (@fig-d4-fig1).
- **QUANT** — calibration fits and concentrations, ending in accuracy against the
  SKML EQA reference values (@fig-d4-fig2).

# Raw Data Processing: Peak Picking and Integration

Raw `.d` files were converted to mzML and integrated with MRMhub-INTEGRATOR, as
described for Datasets 1 and 3. The batch was integrated in two subsets — the
**system-suitability** injections (`SST`, ten replicate injections) and the
**assay** samples (`ASSAY`: blanks, calibrators, QCs, EQA samples) — exported as
separate long-format tables and combined downstream. INTEGRATOR reports peak
areas, retention times and peak metadata in `long.csv`, and writes per-transition
integration plots (the shaded, integrated region) to the `by_transition` folder.

The same raw files were independently re-integrated in Agilent MassHunter
Quantitative Analysis, providing the reference peak areas and concentrations used
throughout this document.

```{r chunk2-integrator-examples}
#| label: fig-d4-chromatograms
#| fig-cap: "**Example INTEGRATOR chromatograms.** MRMhub-INTEGRATOR chromatograms for 21-deoxycortisol, 11-deoxycortisol, Aldosterone, Cortisol D4 (internal standard) and Dihydrotestosterone at high (HQC) and low (LQC) QC levels; the shaded region is the integrated peak."
#| echo: false
#| fig-width: 10
#| fig-height: 4.8
#| fig.alt: "Integrated chromatograms for 21-deoxycortisol, 11-deoxycortisol, Aldosterone, Cortisol D4 and Dihydrotestosterone at HQC and LQC quality-control levels."
# Chromatograms are pre-cropped from the INTEGRATOR by_transition PDFs by
# figures/dataset4/dataset4-chromatograms.R and embedded here as static PNGs, so
# the notebook render needs neither the PDFs nor magick/ghostscript.
chrom_dir <- "images/dataset4"
chrom_grob <- function(name) {
  grid::rasterGrob(
    png::readPNG(file.path(chrom_dir, paste0(name, ".png"))),
    interpolate = TRUE
  )
}

chrom_gg <- function(name, title, subtitle = NULL) {
  ggplot() +
    annotation_custom(chrom_grob(name)) +
    labs(title = title, subtitle = subtitle) +
    theme_void(base_size = 8, base_family = "sans") +
    theme(
      plot.title = element_text(
        size = 8,
        hjust = 0.5, margin = margin(b = 1)
      ),
      plot.subtitle = element_text(
        size = 6.8,
        hjust = 0.5, colour = "grey30", margin = margin(b = 2)
      )
    )
}

(chrom_gg("top_21-deoxycortisol_HQC", "21-deoxycortisol", "HQC") |
  chrom_gg("top_11-deoxycortisol_HQC", "11-deoxycortisol", "HQC") |
  chrom_gg("top_Aldosterone_HQC", "Aldosterone", "HQC") |
  chrom_gg("top_CortisolD4_HQC", "Cortisol D4", "HQC") |
  chrom_gg("top_Dihydrotestosterone_HQC", "Dihydrotestosterone", "HQC")) /
  (chrom_gg("top_21-deoxycortisol_LQC", NULL, "LQC") |
    chrom_gg("top_11-deoxycortisol_LQC", NULL, "LQC") |
    chrom_gg("top_Aldosterone_LQC", NULL, "LQC") |
    chrom_gg("top_CortisolD4_LQC", NULL, "LQC") |
    chrom_gg("top_Dihydrotestosterone_LQC", NULL, "LQC"))
```

::: {.callout-note}
The example chromatograms are cropped from the INTEGRATOR `by_transition` PDFs by
`figures/dataset4/dataset4-chromatograms.R` (run once; requires the `magick` R
package and **ghostscript**) and embedded here as static images from
`images/dataset4/`. The notebook render itself needs neither.
:::

# Data Postprocessing and QC

Unlike Datasets 1 and 3, this is a short pipeline: a 15-analyte, single-batch panel
needs no drift/batch correction or feature filtering. The QUANT module
(`library(mrmhub)`) imports the INTEGRATOR results and metadata, normalizes to the
internal standards, fits external calibration curves, and quantifies.

::: callout-note
## mrmhub version
This notebook was rendered with **mrmhub 0.9.8** (QUANT module of
[MRMhub](https://github.com/SLINGhub/MRMhub)). This API is currently on the
`development` branch; install with
`remotes::install_github("SLINGhub/MRMhub", ref = "development")`. The coloured
console output additionally uses the `fansi` package (`install.packages("fansi")`).
:::

```{r chunk3-loadpkg}
#| cache: false
library(mrmhub)
library(mirai) # parallel processing (picked up automatically by mrmhub)

# Colour mrmhub's cli console feedback in the rendered HTML (requires the fansi package)
mrmhub_enable_cli_color()

# Set number of cores for parallel processing (all-but-one, so the system stays responsive)
n_cores <- {
  n <- parallel::detectCores()
  if (is.na(n)) n <- 4L
  max(1L, n - 1L)
}
if (mirai::status()$daemons == 0) mirai::daemons(n_cores)
```

## Import INTEGRATOR results and metadata

The two INTEGRATOR subsets share columns and hold disjoint analyses, so they are
combined by row-binding. The combined table is written to `output/` because it is
re-imported later (@fig-d4-fig2) with MassHunter areas substituted in.

```{r chunk4-combine-import}
#| cache: false
sst_long <- read_csv(
  file.path(d4, "Dataset4_MRMhub-INTEGRATOR_SST.csv"),
  show_col_types = FALSE
)
assay_long <- read_csv(
  file.path(d4, "Dataset4_MRMhub-INTEGRATOR_ASSAY.csv"),
  show_col_types = FALSE
)
long_combined <- bind_rows(sst_long, assay_long)

long_combined_path <- file.path("output", "Dataset4_long_combined.csv")
write_csv(long_combined, long_combined_path)

# Analysis ids per subset (import strips the ".mzML" extension).
sst_analysis_ids <- unique(str_remove(
  sst_long$raw_data_filename,
  "[.]mzML$"
))
assay_analysis_ids <- unique(str_remove(
  assay_long$raw_data_filename,
  "[.]mzML$"
))

meta_path <- file.path(d4, "Dataset4_Metadata.xlsx")

mexp <- MRMhubExperiment(title = "Dataset 4 — Steroid Assay")
mexp <- import_data_mrmhub(
  mexp,
  path = long_combined_path,
  import_metadata = TRUE
)
mexp <- import_metadata_msorganiser(
  mexp,
  path = meta_path,
  excl_unmatched_analyses = TRUE, ignore_warnings = TRUE
)
```

## Normalization and calibration

Each feature is normalized to its stable-isotope-labelled internal standard, then
a quadratic, `1/x`-weighted external calibration curve is fitted per analyte from
the `Cal 0`–`Cal F` levels.

```{r chunk5-normalize-calibrate}
mexp <- normalize_by_istd(mexp)
mexp <- calc_calibration_results(
  mexp,
  fit_overwrite = TRUE,
  fit_model = "quadratic",
  fit_weighting = "1/x"
)
```

The fitted curves and their metrics are inspected before quantification.

```{r chunk6-calibration-curves}
#| label: fig-d4-calibration-curves
#| fig-cap: "**External calibration curves.** Quadratic, 1/x-weighted curves for the 15 steroid analytes (quantifier transitions)."
#| fig-width: 9
#| fig-height: 7
#| fig.alt: "Calibration curves for the 15 steroid analytes."
plot_calibrationcurves(
  mexp,
  fit_overwrite = FALSE,
  fit_model = "quadratic",
  fit_weighting = "1/x",
  include_istd = FALSE,
  include_qualifier = FALSE,
  show_progress = FALSE
)
```

```{r chunk7-calibration-metrics}
#| label: tbl-d4-calibration-metrics
#| tbl-cap: "**Calibration metrics per analyte.** Fit model, weighting, R², LOD and LOQ for each quantifier."
get_calibration_metrics(mexp) |>
  filter(is_quantifier) |>
  transmute(
    analyte = feature_id,
    fit_model, fit_weighting, r2 = round(r2, 5), lod = signif(lod, 3),
    loq = signif(loq, 3)
  )
```

## Quantification

Concentrations are computed from the normalized intensities via the calibration
curves. Qualifier transitions are excluded; analytes whose curve failed to fit
are skipped rather than aborting the run.

```{r chunk8-quantify}
mexp <- quantify_by_calibration(
  mexp,
  fit_overwrite = FALSE,
  include_qualifier = FALSE,
  ignore_failed_calibration = TRUE,
  fit_model = "quadratic",
  fit_weighting = "1/x"
)

# Export + reload the concentration table (wide: analysis_id, qc_type, analytes).
conc_path <- file.path("output", "Dataset4_quant_results.csv")
save_dataset_csv(
  mexp,
  path = conc_path,
  variable = "conc",
  add_qctype = TRUE
)
d_conc_raw <- read_csv(conc_path, show_col_types = FALSE)
```

## MassHunter re-integration

The MassHunter Quantitative Analysis export is imported as a second experiment
and reshaped to one tidy table — the single MassHunter source for all
comparisons below. The trailing ` (ISTD)` suffix on labelled internal standards
is stripped so feature identifiers match MRMhub.

```{r chunk9-masshunter}
mh_reint <- import_data_masshunter(
  MRMhubExperiment(title = "MassHunter re-integration"),
  path = file.path(
    d4,
    "Dataset4_IntegrationResults_MassHunter_IsoSP-22_V3.csv"
  ),
  import_metadata = TRUE, silent = TRUE
)
mh_data <- get_analyticaldata(mh_reint, annotated = FALSE) |>
  transmute(
    analysis_id,
    feature_id = str_remove(feature_id, "[ ]*\\(ISTD\\)$"),
    rt_mh = feature_rt, area_mh = feature_area,
    conc_mh = feature_conc_final, mi_mh = feature_manual_integration
  )
```

## Below-Cal-A exclusion for study samples

Study samples (`SPL` = the SKML EQA samples) whose **quantifier peak area falls
below the corresponding `Cal A` area** are below the lowest non-zero calibrator
and treated as not reliably quantifiable. The rule is applied per analyte and
**independently per method** — MRMhub SPL area vs MRMhub Cal A area, MassHunter
SPL area vs MassHunter Cal A area — yielding the flags `excl_mrm` / `excl_mh` that
downstream comparisons apply to their SPL rows. Internal standards are never
filtered.

```{r chunk10-spl-exclusion}
CAL_A_ID <- "Cal A_P1"

# MRMhub quantifier areas (quantifiers only, excluding the ISTDs).
mrm_q_area <- get_analyticaldata(mexp, annotated = TRUE) |>
  filter(is_quantifier, !is_istd) |>
  transmute(
    analysis_id,
    analyte_id = feature_id, qc_type, area_mrm = feature_area
  )

cala_mrm <- mrm_q_area |>
  filter(analysis_id == CAL_A_ID) |>
  select(analyte_id, cala_area_mrm = area_mrm)
cala_mh <- mh_data |>
  filter(analysis_id == CAL_A_ID) |>
  transmute(analyte_id = feature_id, cala_area_mh = area_mh)

# One row per SPL (analysis, analyte); flag below-Cal-A per method (NA-safe).
spl_area_excl <- mrm_q_area |>
  filter(qc_type == "SPL") |>
  left_join(
    select(mh_data, analysis_id, analyte_id = feature_id, area_mh),
    by = c("analysis_id", "analyte_id")
  ) |>
  left_join(cala_mrm, by = "analyte_id") |>
  left_join(cala_mh, by = "analyte_id") |>
  mutate(
    excl_mrm = !is.na(area_mrm) & !is.na(cala_area_mrm) &
      area_mrm < cala_area_mrm,
    excl_mh = !is.na(area_mh) & !is.na(cala_area_mh) &
      area_mh < cala_area_mh
  ) |>
  select(
    analysis_id, analyte_id,
    area_mrm, area_mh, cala_area_mrm, cala_area_mh,
    excl_mrm, excl_mh
  )
```

## Low / high QC

Quantified concentrations of the low and high QC samples, one row per analyte.

```{r chunk11-qc-highlow}
#| label: tbl-d4-qc-conc
#| tbl-cap: "**Low/high QC concentrations.** Quantified concentrations of the low and high QC samples, one row per analyte."
d_conc_raw |>
  filter(qc_type %in% c("LQC", "HQC")) |>
  pivot_longer(
    !c(analysis_id, qc_type),
    names_to = "analyte", values_to = "conc"
  ) |>
  mutate(qc = factor(qc_type, levels = c("LQC", "HQC"))) |>
  select(analyte, qc, conc) |>
  pivot_wider(names_from = qc, values_from = conc) |>
  mutate(across(where(is.numeric), ~ signif(.x, 3)))
```

# MRMhub vs MassHunter — Peak Integration

@fig-d4-fig1 compares MRMhub-INTEGRATOR peak areas with the MassHunter
re-integration of the same raw files. MRMhub quantifier and ISTD peaks are joined
to MassHunter on `analysis_id` + `feature_id`, dropping near-noise peaks
(area < 100), empty CAL rows, and below-Cal-A SPL points.

```{r chunk12-house-style}
# Manuscript house theme + palettes (shared by both figures).
theme_set(
  theme_bw(base_size = 8, base_family = "sans") +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      plot.title = element_text(size = 8, face = "plain"),
      legend.key.size = unit(3.2, "mm"), legend.margin = margin(1, 1, 1, 1)
    )
)

col_method <- c(MRMhub = "#2C7FB8", MassHunter = "#D95F0E")
smp_cols <- c(
  CAL = "#0072B2",
  HQC = "#D55E00", LQC = "#E69F00", SPL = "grey30", SST = "#009E73",
  SBLK = "#CC79A7", PBLK = "#999999"
)

# Save a composite as vector PDF + 300-dpi PNG at a given size (mm) for the manuscript.
save_fig <- function(plot, stem, height, width = 180) {
  ggsave(
    file.path("output", paste0(stem, ".pdf")),
    plot,
    device = cairo_pdf, width = width, height = height, units = "mm"
  )
  ggsave(
    file.path("output", paste0(stem, ".png")),
    plot,
    device = ragg::agg_png, width = width, height = height,
    units = "mm", dpi = 300
  )
  invisible(plot)
}
```

```{r chunk13-fig1-data}
mrm_peaks_q <- get_analyticaldata(mexp, annotated = TRUE) |>
  filter(is_quantifier | is_istd) |>
  transmute(
    analysis_id,
    feature_id, qc_type, sample_id,
    feature_type = factor(
      if_else(is_istd, "ISTD", "Quantifier"),
      levels = c("Quantifier", "ISTD")
    ),
    rt_mrm = feature_rt, area_mrm = feature_area
  )

d_peakcmp_reint <- mrm_peaks_q |>
  inner_join(mh_data, by = c("analysis_id", "feature_id")) |>
  filter(
    !is.na(rt_mrm),
    !is.na(rt_mh), !is.na(area_mrm), !is.na(area_mh)
  ) |>
  filter(area_mrm >= 100, area_mh >= 100) |>
  filter(!(qc_type == "CAL" & is.na(sample_id) & feature_type != "ISTD")) |>
  left_join(
    select(spl_area_excl, analysis_id, analyte_id, excl_mrm, excl_mh),
    by = c("analysis_id", "feature_id" = "analyte_id")
  ) |>
  filter(
    !(qc_type == "SPL" &
      (coalesce(excl_mrm, FALSE) | coalesce(excl_mh, FALSE)))
  ) |>
  select(-excl_mrm, -excl_mh) |>
  mutate(
    area_diff_pct = (area_mh - area_mrm) / area_mrm * 100,
    rt_diff = rt_mh - rt_mrm,
    subset = case_when(
      analysis_id %in% sst_analysis_ids ~ "SST",
      analysis_id %in% assay_analysis_ids ~ "ASSAY", TRUE ~ NA_character_
    )
  )
```

## Peak-area agreement

Log–log peak areas against the 1:1 line (Pearson *r* on log10 values). CAL/QC/SST
samples span four orders of magnitude, and the two integrators agree closely.

```{r chunk14-fig1-panelA}
#| label: fig-d4-fig1-panelA
#| fig-cap: "**Peak-area agreement.** MRMhub vs MassHunter peak areas (log–log) against the 1:1 line; Pearson *r* on log10 values across CAL/QC/SST samples. Open grey rings mark peaks MassHunter integrated manually. Panel A of @fig-d4-fig1."
#| fig-width: 4.6
#| fig-height: 4.2
smp_lv <- c("CAL", "HQC", "LQC", "SST")
smp_cols4 <- c(
  CAL = "#0072B2",
  HQC = "#D55E00", LQC = "#E69F00", SST = "#009E73"
)
ft_shapes <- c(Quantifier = 16, ISTD = 17)
scale_smp <- scale_colour_manual(
  values = smp_cols4,
  limits = smp_lv, drop = FALSE
)
scale_ft <- scale_shape_manual(values = ft_shapes)

d_cmp <- d_peakcmp_reint |>
  filter(qc_type %in% smp_lv) |>
  mutate(qc_type = factor(qc_type, levels = smp_lv))
r_log <- cor(log10(d_cmp$area_mrm), log10(d_cmp$area_mh))
ax_lims <- range(c(d_cmp$area_mrm, d_cmp$area_mh))
p_corr <- d_cmp |>
  ggplot(aes(area_mrm, area_mh)) +
  geom_abline(
    slope = 1,
    intercept = 0, linetype = "dashed", colour = "grey40"
  ) +
  geom_point(
    aes(colour = qc_type, shape = feature_type),
    size = 1.1, alpha = 0.7
  ) +
  # Ring the peaks MassHunter integrated manually (MI flag), keeping the
  # sample-type colour / feature shape encoding underneath.
  geom_point(
    data = ~ filter(.x, mi_mh),
    aes(alpha = "Manual integration (MassHunter)"), shape = 1,
    colour = "grey15", size = 2.3, stroke = 0.5
  ) +
  annotate(
    "text",
    x = ax_lims[1], y = ax_lims[2], hjust = 0, vjust = 1, size = 2.4,
    label = sprintf("Pearson r = %.4f", r_log)
  ) +
  scale_x_log10(limits = ax_lims) +
  scale_y_log10(limits = ax_lims) +
  scale_smp +
  scale_ft +
  scale_alpha_manual(
    name = NULL,
    values = c("Manual integration (MassHunter)" = 1)
  ) +
  guides(
    colour = guide_legend(order = 1, nrow = 2),
    shape = guide_legend(order = 2, nrow = 2),
    alpha = guide_legend(
      order = 3,
      override.aes = list(
        shape = 1,
        colour = "grey15", size = 2.3, stroke = 0.5
      )
    )
  ) +
  labs(
    x = "MRMhub peak area",
    y = "MassHunter peak area", colour = "Sample type", shape = "Feature",
    title = "Peak-area agreement"
  )
p_corr
```

## Precision (%CV) equivalence

Per-feature %CV over the ten replicate SST injections, MRMhub vs MassHunter. The
two methods show comparable precision, with points scattering around the 1:1 line.

```{r chunk15-fig1-panelB}
#| label: fig-d4-fig1-panelB
#| fig-cap: "**Precision (%CV) equivalence.** Per-feature %CV over the ten replicate SST injections, MRMhub vs MassHunter; points on the 1:1 line indicate equal precision. Panel B of @fig-d4-fig1."
#| fig-width: 4.6
#| fig-height: 4.2
sst_cv <- d_peakcmp_reint |>
  filter(subset == "SST") |>
  group_by(feature_type, feature_id) |>
  summarise(
    cv_mrmhub = sd(area_mrm) / mean(area_mrm) * 100,
    cv_masshunter = sd(area_mh) / mean(area_mh) * 100, .groups = "drop"
  )
cv_lims <- range(c(sst_cv$cv_mrmhub, sst_cv$cv_masshunter)) + c(-0.3, 0.3)
p_cv <- sst_cv |>
  ggplot(aes(cv_mrmhub, cv_masshunter, shape = feature_type)) +
  geom_abline(
    slope = 1,
    intercept = 0, linetype = "dashed", colour = "grey60"
  ) +
  geom_point(size = 1.6, alpha = 0.85, colour = "#009E73") +
  coord_cartesian(xlim = cv_lims, ylim = cv_lims) +
  scale_ft +
  guides(shape = "none") +
  labs(
    x = "MRMhub %CV (10 SST reps)",
    y = "MassHunter %CV", title = "Precision (%CV) equivalence"
  )
p_cv
```

## Bland–Altman

Percent difference (MassHunter − MRMhub) against the mean area, excluding SST and
near-noise peaks (mean area < 10³). The three highlighted analytes are among those
with the largest real-sample differences (see @tbl-d4-chrom-candidates); their
chromatograms are shown below.

```{r chunk16-fig1-panelC}
#| label: fig-d4-fig1-panelC
#| fig-cap: "**Bland–Altman of real-sample differences.** Percent difference (MassHunter − MRMhub) against the mean peak area (excl. SST; mean area ≥ 10³); the three largest-difference analytes are highlighted and open grey rings mark manual MassHunter integrations. Panel C of @fig-d4-fig1."
#| fig-width: 7
#| fig-height: 3.4
highlight_analytes <- c(
  "21-deoxycortisol",
  "11-deoxycortisol", "Aldosterone"
)
hl_cols <- setNames(c("#D81B60", "#1E88E5", "#8E44AD"), highlight_analytes)

d_ba <- d_peakcmp_reint |>
  filter(qc_type != "SST") |>
  mutate(
    mean_area = (area_mrm + area_mh) / 2,
    pct_diff = (area_mh - area_mrm) / mean_area * 100,
    hl = factor(
      if_else(
        feature_id %in% highlight_analytes,
        feature_id, NA_character_
      ),
      levels = highlight_analytes
    )
  ) |>
  filter(mean_area >= 1e3)
ba_bias <- mean(d_ba$pct_diff)
ba_loa <- ba_bias + c(-1.96, 1.96) * sd(d_ba$pct_diff)
ba_xr <- max(d_ba$mean_area)
p_ba <- d_ba |>
  ggplot(aes(mean_area, pct_diff)) +
  geom_hline(yintercept = 0, colour = "grey70") +
  geom_hline(yintercept = ba_bias, colour = "grey20") +
  geom_hline(yintercept = ba_loa, linetype = "dashed", colour = "grey50") +
  geom_point(
    data = ~ filter(.x, is.na(hl)),
    colour = "grey80", size = 0.9, alpha = 0.6
  ) +
  geom_point(
    data = ~ filter(.x, !is.na(hl)),
    aes(colour = hl), size = 1.7, alpha = 0.9
  ) +
  # Ring the peaks MassHunter integrated manually (MI flag).
  geom_point(
    data = ~ filter(.x, mi_mh),
    aes(alpha = "Manual integration (MassHunter)"), shape = 1,
    colour = "grey15", size = 2.0, stroke = 0.5
  ) +
  annotate(
    "text",
    x = ba_xr, y = ba_bias, hjust = 1, vjust = -0.6, size = 2.1,
    colour = "grey20", label = sprintf("mean bias %.1f%%", ba_bias)
  ) +
  annotate(
    "text",
    x = ba_xr, y = ba_loa[2], hjust = 1, vjust = 1.4, size = 2,
    colour = "grey45", label = "+95% LoA"
  ) +
  annotate(
    "text",
    x = ba_xr, y = ba_loa[1], hjust = 1, vjust = -0.7, size = 2,
    colour = "grey45", label = "-95% LoA"
  ) +
  scale_colour_manual(
    values = hl_cols,
    name = "Highlighted analyte", na.translate = FALSE
  ) +
  scale_alpha_manual(
    name = NULL,
    values = c("Manual integration (MassHunter)" = 1)
  ) +
  scale_x_log10(limits = c(1e3, NA)) +
  coord_cartesian(ylim = c(-20, 20)) +
  guides(
    alpha = guide_legend(
      order = 2,
      override.aes = list(
        shape = 1,
        colour = "grey15", size = 2.0, stroke = 0.5
      )
    )
  ) +
  labs(
    x = "Mean peak area of the two methods (>= 10^3)",
    y = "Percent difference (MH - MRMhub) %",
    title = sprintf(
      "Bland-Altman (excl. SST, area >= 10^3): mean bias %.1f%%, 95%% LoA [%.0f, %.0f]%%",
      ba_bias, ba_loa[1], ba_loa[2]
    )
  )
p_ba
```

## Example chromatograms

The largest real-sample area differences occur where a peak sits on an elevated
baseline or beside a co-eluting neighbour, so the integration boundaries — and
hence the areas — legitimately differ between the two integrators. The candidate
table ranks these cases; the panel crops the three chosen examples (one per
highlighted analyte) from the MRMhub `by_transition` chromatograms.

```{r chunk17-fig1-chrom-candidates}
#| label: tbl-d4-chrom-candidates
#| tbl-cap: "**Largest area-difference candidates.** Peaks ranked by absolute MRMhub–MassHunter area difference, from which the example chromatograms are chosen."
chrom_candidates <- d_peakcmp_reint |>
  filter(
    feature_type == "Quantifier",
    qc_type %in% c("CAL", "HQC", "LQC", "SPL", "SST"), area_mrm >= 20000
  ) |>
  arrange(desc(abs(area_diff_pct))) |>
  transmute(
    feature_id,
    analysis_id, qc_type,
    area_mrm = round(area_mrm),
    area_mh = round(area_mh), area_diff_pct = round(area_diff_pct, 1)
  ) |>
  head(10)
chrom_candidates
```

```{r chunk18-fig1-chrom-panel}
#| label: fig-d4-fig1-panelD
#| fig-cap: "**Example chromatograms.** MRMhub chromatograms for the three highlighted analytes, where the integration boundaries — and hence the areas — legitimately differ between the two integrators. Panel D of @fig-d4-fig1."
#| fig-width: 8
#| fig-height: 3
# Panel-D chromatograms are the pre-generated crops (see the INTEGRATOR section);
# subtitles carry the data-derived area difference for each example.
chrom_examples <- tibble::tribble(
  ~name,
  ~feature_id, ~analysis_id, "fig1d_21-deoxycortisol", "21-deoxycortisol",
  "SKML2025_6B_P1", "fig1d_11-deoxycortisol", "11-deoxycortisol",
  "SKML2026_2B_P1", "fig1d_Aldosterone", "Aldosterone", "SKML2026_2A_P1"
)

chrom_meta <- chrom_examples |>
  left_join(
    distinct(d_peakcmp_reint, feature_id, analysis_id, area_diff_pct),
    by = c("feature_id", "analysis_id")
  ) |>
  mutate(
    subtitle = sprintf(
      "%s  (dArea %+.1f%%)",
      str_remove(analysis_id, "_P1$"), area_diff_pct
    )
  )
chrom_plots <- Map(
  chrom_gg,
  chrom_meta$name, chrom_meta$feature_id, chrom_meta$subtitle
)
p_chrom <- wrap_elements(full = wrap_plots(chrom_plots, nrow = 1))
p_chrom
```

## Figure 1 — manuscript figure

The four panels above are composed into the publication figure (**Figure 1**) and
saved as a vector PDF and a 300-dpi PNG (180 mm double-column width).

```{r chunk19-fig1-assemble}
#| label: fig-d4-fig1
#| fig-cap: "**MRMhub vs MassHunter — peak integration.** Composite of the four panels above; reproduced as the manuscript's **Figure 1**. **(A)** Peak-area agreement, **(B)** precision (%CV) equivalence, **(C)** Bland–Altman of real-sample differences, **(D)** example chromatograms. Exported to `output/fig1_integration_correspondence.{pdf,png}` (180 mm)."
#| fig-width: 7.1
#| fig-height: 9.06
row_ab <- (p_corr + p_cv) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
fig1 <- row_ab /
  p_ba /
  p_chrom +
  plot_layout(heights = c(1, 0.72, 0.58)) +
  plot_annotation(tag_levels = "A")

save_fig(fig1, "fig1_integration_correspondence", height = 230)
fig1
```

# MRMhub vs MassHunter — Quantification

@fig-d4-fig2 compares the two pipelines at the quantification stage. To isolate the
post-processing (calibration + quantification) from the peak integration, we also
feed **MassHunter's own peak areas through the MRMhub pipeline** — identical
areas, identical calibration model — so any concentration difference is due to
post-processing alone.

```{r chunk20-fig2-mhpipe}
mh_area <- mh_data |>
  transmute(analysis_id, feature_name = feature_id, area_mh)
long_mh <- read_csv(long_combined_path, show_col_types = FALSE) |>
  mutate(.aid = str_remove(raw_data_filename, "[.]mzML$")) |>
  inner_join(mh_area, by = c(".aid" = "analysis_id", "feature_name")) |>
  mutate(area = area_mh) |>
  select(-.aid, -area_mh)

tmp_long <- tempfile(fileext = ".csv")
write_csv(long_mh, tmp_long)

mexp_mh <- MRMhubExperiment(title = "MassHunter areas via MRMhub pipeline")
mexp_mh <- import_data_mrmhub(
  mexp_mh,
  path = tmp_long,
  import_metadata = TRUE
)
mexp_mh <- import_metadata_msorganiser(
  mexp_mh,
  path = meta_path,
  excl_unmatched_analyses = TRUE, ignore_warnings = TRUE
)
mexp_mh <- normalize_by_istd(mexp_mh)
mexp_mh <- calc_calibration_results(
  mexp_mh,
  fit_overwrite = TRUE,
  fit_model = "quadratic",
  fit_weighting = "1/x"
)
mexp_mh <- quantify_by_calibration(
  mexp_mh,
  fit_overwrite = FALSE,
  include_qualifier = FALSE,
  ignore_failed_calibration = TRUE,
  fit_model = "quadratic",
  fit_weighting = "1/x"
)
```

## Calibration fit comparison

Calibration R² per analyte: MassHunter's own fit vs the MRMhub full-pipeline fit.

```{r chunk21-fig2-panelA}
#| label: fig-d4-fig2-panelA
#| fig-cap: "**Calibration-fit comparison.** R² per analyte, MassHunter's own fit vs the MRMhub full-pipeline fit. Panel A of @fig-d4-fig2."
#| fig-width: 6
#| fig-height: 4.5
mh_own_r2 <- read_csv(
  file.path(d4, "Dataset4_MassHunter-CalibrationFits.csv"),
  show_col_types = FALSE
) |>
  transmute(analyte_id, r2_mh_own = r2)
r2_cmp <- get_calibration_metrics(mexp_mh) |>
  filter(is_quantifier) |>
  transmute(analyte_id = feature_id, r2_mrmpipe_mharea = r2) |>
  left_join(mh_own_r2, by = "analyte_id") |>
  left_join(
    get_calibration_metrics(mexp) |>
      filter(is_quantifier) |>
      transmute(analyte_id = feature_id, r2_mrmpipe_mrmarea = r2),
    by = "analyte_id"
  )

chain_lv <- c("MassHunter", "MRMhub")
chain_col <- c(MassHunter = "#D95F0E", MRMhub = "#2C7FB8")
ord_r2 <- r2_cmp |>
  arrange(r2_mrmpipe_mrmarea) |>
  pull(analyte_id)

d_r2_long <- r2_cmp |>
  transmute(
    analyte_id,
    MassHunter = r2_mh_own, MRMhub = r2_mrmpipe_mrmarea
  ) |>
  pivot_longer(-analyte_id, names_to = "chain", values_to = "r2") |>
  mutate(
    analyte_id = factor(analyte_id, levels = ord_r2),
    chain = factor(chain, levels = chain_lv)
  )

# Truncate the x-axis just below the dense 0.996–1.000 cluster; flag off-scale rows.
R2_FLOOR <- 0.9955
r2_off <- d_r2_long |>
  filter(r2 < R2_FLOOR) |>
  distinct(analyte_id) |>
  mutate(r2 = R2_FLOOR)
off_txt <- d_r2_long |>
  filter(r2 < R2_FLOOR) |>
  group_by(analyte_id) |>
  summarise(v = min(r2), .groups = "drop") |>
  arrange(v) |>
  mutate(t = sprintf("%s %.3f", analyte_id, v))
p_r2 <- ggplot(d_r2_long, aes(r2, analyte_id)) +
  geom_line(aes(group = analyte_id), colour = "grey75", linewidth = 0.6) +
  geom_point(aes(colour = chain), size = 1.9, alpha = 0.9) +
  geom_text(
    data = r2_off,
    aes(label = "<"), hjust = 0, nudge_x = 0.00004, size = 2.3,
    colour = "grey45"
  ) +
  scale_colour_manual(values = chain_col, name = NULL) +
  scale_x_continuous(breaks = c(0.996, 0.997, 0.998, 0.999, 1.000)) +
  coord_cartesian(xlim = c(R2_FLOOR, 1.0004)) +
  labs(
    x = expression(Calibration ~ R^2),
    y = NULL, title = "Calibration curve fit comparison",
    subtitle = paste0(
      "x-axis truncated at ",
      R2_FLOOR, "; off scale (<): ", paste(off_txt$t, collapse = ", ")
    )
  ) +
  theme(
    plot.subtitle = element_text(size = 6.5, colour = "grey30"),
    plot.margin = margin(4, 10, 4, 4), legend.position = "inside",
    legend.position.inside = c(0.015, 0.98),
    legend.justification.inside = c(0, 1),
    legend.background = element_rect(
      fill = alpha("white", 0.7),
      colour = NA
    ),
    legend.key.size = unit(3, "mm")
  )
p_r2
```

The same comparison as a table, with **three fits per analyte** (quadratic,
`y = a·x² + b·x + c`, 1/x weighting): MassHunter's own fit, the MRMhub end-to-end
fit (MRMhub-INTEGRATOR areas), and the MRMhub pipeline run on **MassHunter's own
peak areas**. The gap between the two MRMhub rows is the peak-integration effect;
the distance between *MRMhub (on MH areas)* and *MassHunter* isolates the
post-processing (normalization, weighting, regression) alone.

```{r chunk21b-calfit-coef-table}
#| label: tbl-d4-calfit-coef
#| tbl-cap: "**Calibration fit per analyte.** Coefficients of y = a·x² + b·x + c, R², and measured LQC/HQC concentrations (nmol/L) for three fits — MassHunter's own, MRMhub on MassHunter areas, and MRMhub end-to-end (blue text). The two MRMhub rows differ only by peak integration; MRMhub-on-MH-areas vs MassHunter isolates post-processing."
library(gt)

# Measured LQC / HQC concentrations (nmol/L) per engine, one column each.
qc_wide <- function(df) {
  df |>
    filter(qc_type %in% c("LQC", "HQC")) |>
    pivot_longer(
      !c(analysis_id, qc_type),
      names_to = "analyte_id", values_to = "conc"
    ) |>
    select(analyte_id, qc_type, conc) |>
    pivot_wider(names_from = qc_type, values_from = conc)
}

tmp_mhconc <- tempfile(fileext = ".csv")
save_dataset_csv(
  mexp_mh,
  path = tmp_mhconc,
  variable = "conc",
  add_qctype = TRUE
)
conc_e2e <- qc_wide(d_conc_raw)
conc_onmh <- qc_wide(read_csv(tmp_mhconc, show_col_types = FALSE))
conc_mhun <- mh_data |>
  filter(analysis_id %in% c("QC_Low1_P1", "QC_High1_P1")) |>
  transmute(
    analyte_id = feature_id,
    qc_type = if_else(analysis_id == "QC_Low1_P1", "LQC", "HQC"),
    conc = conc_mh
  ) |>
  pivot_wider(names_from = qc_type, values_from = conc)

# mrmhub reports coefficients intercept-first (coef_a = x^0 ... coef_c = x^2);
# remap to MassHunter's descending order (a = x^2, b = x, c = intercept).
mh_own <- read_csv(
  file.path(d4, "Dataset4_MassHunter-CalibrationFits.csv"),
  show_col_types = FALSE
) |>
  transmute(
    analyte_id,
    method = "MassHunter", a = coef_a, b = coef_b, c = coef_c, r2 = r2
  ) |>
  left_join(conc_mhun, by = "analyte_id")
mrm_mh <- get_calibration_metrics(mexp_mh) |>
  filter(is_quantifier) |>
  transmute(
    analyte_id = feature_id,
    method = "MRMhub (on MH areas)", a = coef_c, b = coef_b, c = coef_a,
    r2 = r2
  ) |>
  left_join(conc_onmh, by = "analyte_id")
mrm_e2e <- get_calibration_metrics(mexp) |>
  filter(is_quantifier) |>
  transmute(
    analyte_id = feature_id,
    method = "MRMhub (end-to-end)", a = coef_c, b = coef_b, c = coef_a,
    r2 = r2
  ) |>
  left_join(conc_e2e, by = "analyte_id")

cal_lv <- c("MassHunter", "MRMhub (on MH areas)", "MRMhub (end-to-end)")
fmt_conc <- function(x) {
  ifelse(
    is.na(x),
    "—",
    formatC(
      signif(x, 4),
      format = "fg", big.mark = ",", drop0trailing = TRUE
    )
  )
}
cal_coef_tbl <- bind_rows(mh_own, mrm_mh, mrm_e2e) |>
  mutate(method = factor(method, levels = cal_lv)) |>
  arrange(analyte_id, method) |>
  transmute(
    Analyte = if_else(as.character(method) == cal_lv[1], analyte_id, ""),
    Method = as.character(method), `a (x²)` = sprintf("%.4g", a),
    `b (x)` = sprintf("%.4g", b), `c` = sprintf("%.4g", c),
    `R²` = sprintf("%.5f", r2), `LQC Concentration` = fmt_conc(LQC),
    `HQC Concentration` = fmt_conc(HQC)
  )

gt(cal_coef_tbl) |>
  cols_align(
    "right",
    columns = c(
      "a (x²)",
      "b (x)", "c", "R²", "LQC Concentration", "HQC Concentration"
    )
  ) |>
  tab_style(
    cell_fill(color = "#DEEBF7"),
    cells_body(rows = Method == "MRMhub (on MH areas)")
  ) |>
  tab_style(
    list(cell_fill(color = "#2C7FB8"), cell_text(color = "white")),
    cells_body(rows = Method == "MRMhub (end-to-end)")
  ) |>
  tab_options(table.font.size = px(12), data_row.padding = px(2))
```

## Concentration equivalence on identical areas

Feeding MassHunter areas through the MRMhub pipeline reproduces MassHunter's
end-to-end concentrations closely, indicating that the post-processing steps are
effectively equivalent.

```{r chunk22-fig2-panelB}
#| label: fig-d4-fig2-panelB
#| fig-cap: "**Concentration equivalence on identical areas.** MassHunter areas run through the MRMhub pipeline vs MassHunter end-to-end, isolating post-processing from integration. Panel B of @fig-d4-fig2."
#| fig-width: 4.6
#| fig-height: 4.4
tmp_conc <- tempfile(fileext = ".csv")
save_dataset_csv(
  mexp_mh,
  path = tmp_conc,
  variable = "conc",
  add_qctype = TRUE
)
d_qonly <- read_csv(tmp_conc, show_col_types = FALSE) |>
  pivot_longer(
    !c(analysis_id, qc_type),
    names_to = "analyte_id", values_to = "conc_mrmpipe"
  ) |>
  left_join(
    mh_data |> select(analysis_id, analyte_id = feature_id, conc_mh),
    by = c("analysis_id", "analyte_id")
  ) |>
  filter(!is.na(conc_mrmpipe), !is.na(conc_mh)) |>
  filter(!qc_type %in% c("SBLK", "PBLK")) |>
  mutate(qc_type = factor(qc_type, levels = names(smp_cols)))
posq <- d_qonly |> filter(conc_mrmpipe > 0, conc_mh > 0)
q_r <- cor(log10(posq$conc_mrmpipe), log10(posq$conc_mh))
axq <- range(c(posq$conc_mrmpipe, posq$conc_mh))
p_qonly <- posq |>
  ggplot(aes(conc_mh, conc_mrmpipe, colour = qc_type)) +
  geom_abline(
    slope = 1,
    intercept = 0, linetype = "dashed", colour = "grey40"
  ) +
  geom_point(size = 1.1, alpha = 0.6) +
  annotate(
    "text",
    x = axq[1], y = axq[2], hjust = 0, vjust = 1, size = 2.4,
    label = sprintf("Pearson r = %.5f", q_r)
  ) +
  scale_x_log10(limits = axq) +
  scale_y_log10(limits = axq) +
  scale_colour_manual(
    values = smp_cols,
    drop = TRUE, name = "Sample type"
  ) +
  coord_equal() +
  labs(
    x = "MassHunter end-to-end (nmol/L)",
    y = "MRMhub-QUANT on MH areas (nmol/L)",
    title = "Concentration equivalence (same areas)"
  ) +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.98, 0.02),
    legend.justification.inside = c(1, 0),
    legend.background = element_rect(
      fill = alpha("white", 0.7),
      colour = NA
    ),
    legend.key.size = unit(2.8, "mm"),
    legend.title = element_text(size = 6.5),
    legend.text = element_text(size = 6)
  )
p_qonly
```

## Full end-to-end concentration agreement

Independent end-to-end concentrations (each method's own integration and
calibration) for the LQC, HQC and SPL samples.

```{r chunk23-fig2-panelC}
#| label: fig-d4-fig2-panelC
#| fig-cap: "**Full end-to-end concentration agreement.** Each method's own integration and calibration, for the LQC, HQC and SPL samples. Panel C of @fig-d4-fig2."
#| fig-width: 4.6
#| fig-height: 4.4
d_full <- d_conc_raw |>
  filter(qc_type %in% c("LQC", "HQC", "SPL")) |>
  pivot_longer(
    !c("analysis_id", "qc_type"),
    names_to = "analyte_id", values_to = "conc"
  ) |>
  left_join(
    mh_data |> select(analysis_id, analyte_id = feature_id, conc_mh),
    by = c("analysis_id", "analyte_id")
  ) |>
  left_join(
    select(spl_area_excl, analysis_id, analyte_id, excl_mrm, excl_mh),
    by = c("analysis_id", "analyte_id")
  ) |>
  mutate(
    conc = if_else(coalesce(excl_mrm, FALSE), NA_real_, conc),
    conc_mh = if_else(coalesce(excl_mh, FALSE), NA_real_, conc_mh)
  ) |>
  select(-excl_mrm, -excl_mh) |>
  filter(!is.na(conc), !is.na(conc_mh), conc > 0, conc_mh > 0) |>
  mutate(qc_type = factor(qc_type, levels = names(smp_cols)))
f_r <- cor(log10(d_full$conc), log10(d_full$conc_mh))
axf <- range(c(d_full$conc, d_full$conc_mh))
p_full <- d_full |>
  ggplot(aes(conc_mh, conc, colour = qc_type)) +
  geom_abline(
    slope = 1,
    intercept = 0, linetype = "dashed", colour = "grey40"
  ) +
  geom_point(size = 1.3, alpha = 0.7) +
  annotate(
    "text",
    x = axf[1], y = axf[2], hjust = 0, vjust = 1, size = 2.4,
    label = sprintf("Pearson r = %.4f", f_r)
  ) +
  scale_x_log10(limits = axf) +
  scale_y_log10(limits = axf) +
  scale_colour_manual(
    values = smp_cols,
    drop = TRUE, name = "Sample type"
  ) +
  coord_equal() +
  labs(
    x = "MassHunter end-to-end (nmol/L)",
    y = "MRMhub end-to-end (nmol/L)",
    title = "Concentration agreement (full pipeline)"
  ) +
  theme(legend.position = "none")
p_full
```

## Accuracy against reference values (QC and EQA)

Per-analyte percent bias against the reference concentrations — the SKML EQA
reference values and the nominal LQC/HQC target concentrations — one point per
sample per method (dashed lines: ±15 %; LQC and HQC are grouped to the left of the
EQA samples). EQA points whose quantifier area falls below the lowest calibrator
(Cal A, i.e. below the limit of quantification) are drawn as open symbols rather
than excluded. The reference concentrations are read from the metadata workbook's
`QC Concentrations` sheet (imported above with the rest of the metadata), keyed by
sample and analyte. DHEA and DHEAS are excluded (large biases that would compress
the other panels).

```{r chunk24-fig2-panelD}
#| label: fig-d4-fig2-panelD
#| fig-cap: "**Accuracy against reference values.** Per-analyte percent bias against the reference concentrations — SKML EQA values plus the nominal LQC/HQC targets (±15 % dashed) — MRMhub vs MassHunter, one point per sample per method, with LQC/HQC grouped to the left of the EQA samples; open symbols mark EQA values below the lowest calibrator (Cal A, below LOQ) and DHEA and DHEAS are excluded. Panel D of @fig-d4-fig2."
#| fig-width: 6.5
#| fig-height: 10
# Reference (target) concentrations come from the metadata's `QC Concentrations`
# sheet (imported above), keyed by `sample_id` + `analyte_id`, and surface via the
# `annot_qcconcentrations` slot (all in nmol/L). Both the SKML EQA reference values
# and the nominal LQC/HQC targets are kept, so accuracy is assessed for all three.
acc_target <- mexp@annot_qcconcentrations |>
  filter(str_starts(sample_id, "SKML") | sample_id %in% c("LQC", "HQC")) |>
  transmute(sample_id, analyte_id, target = concentration)
# analysis_id -> sample_id map (groups the `…b` replicates onto one EQA sample).
sid_map <- get_analyticaldata(mexp, annotated = TRUE) |>
  distinct(analysis_id, sample_id)
d_acc <- d_conc_raw |>
  filter(qc_type %in% c("LQC", "HQC", "SPL")) |>
  pivot_longer(
    !c("analysis_id", "qc_type"),
    names_to = "analyte_id", values_to = "conc_mrm"
  ) |>
  left_join(
    mh_data |> select(analysis_id, analyte_id = feature_id, conc_mh),
    by = c("analysis_id", "analyte_id")
  ) |>
  left_join(
    select(spl_area_excl, analysis_id, analyte_id, excl_mrm, excl_mh),
    by = c("analysis_id", "analyte_id")
  ) |>
  left_join(sid_map, by = "analysis_id") |>
  inner_join(acc_target, by = c("sample_id", "analyte_id")) |>
  # Compact axis label: SKML EQA -> 2026.1A; LQC/HQC keep their name.
  mutate(
    sample_lbl = str_replace(
      str_remove(sample_id, "^SKML"),
      "^([0-9]{4})_", "\\1."
    )
  ) |>
  # Below-Cal-A (below-LOQ) values are kept rather than masked; a per-method flag
  # (`blo_*`) lets the panel draw them as open symbols. LQC/HQC carry no exclusion
  # flag (never below-Cal-A here), so they count as quantifiable.
  transmute(
    analyte_id,
    sample_lbl, analysis_id, target, conc_mrm, conc_mh,
    bias_mrmhub = (conc_mrm - target) / target * 100,
    bias_masshunter = (conc_mh - target) / target * 100,
    blo_mrmhub = coalesce(excl_mrm, FALSE),
    blo_masshunter = coalesce(excl_mh, FALSE)
  ) |>
  filter(!(is.na(conc_mrm) & is.na(conc_mh)))

# x-axis order: LQC, HQC first, then the EQA samples (sorted by label).
eqa_lbls <- sort(setdiff(unique(d_acc$sample_lbl), c("LQC", "HQC")))
x_levels <- c("LQC", "HQC", eqa_lbls)
p_eqa <- d_acc |>
  filter(!analyte_id %in% c("DHEA", "DHEAS")) |>
  # Reshape bias and the below-LOQ flag together (one row per method).
  pivot_longer(
    c(bias_mrmhub, bias_masshunter, blo_mrmhub, blo_masshunter),
    names_to = c(".value", "method"),
    names_pattern = "(bias|blo)_(mrmhub|masshunter)"
  ) |>
  filter(!is.na(bias)) |>
  mutate(
    method = recode(method, mrmhub = "MRMhub", masshunter = "MassHunter"),
    method = factor(method, levels = c("MRMhub", "MassHunter")),
    sample_lbl = factor(sample_lbl, levels = x_levels),
    loq = factor(
      if_else(blo, "Below LOQ (< Cal A)", "Quantifiable (>= Cal A)"),
      levels = c("Quantifiable (>= Cal A)", "Below LOQ (< Cal A)")
    )
  ) |>
  ggplot(aes(
    sample_lbl, bias,
    colour = method, shape = loq, group = method
  )) +
  geom_hline(yintercept = 0, colour = "grey40") +
  geom_hline(
    yintercept = c(-15, 15),
    linetype = "dashed", colour = "grey75"
  ) +
  # Divider between the LQC/HQC group (left) and the EQA samples (right).
  geom_vline(xintercept = 2.5, colour = "grey85", linewidth = 0.3) +
  geom_point(
    size = 1.5,
    alpha = 0.85, position = position_dodge(width = 0.4)
  ) +
  scale_colour_manual(values = col_method, name = "Method") +
  # Open symbols flag EQA values below the lowest calibrator (below LOQ).
  scale_shape_manual(
    values = c("Quantifiable (>= Cal A)" = 16, "Below LOQ (< Cal A)" = 1),
    name = NULL
  ) +
  facet_wrap(vars(analyte_id), scales = "free_y", ncol = 3) +
  guides(
    colour = guide_legend(override.aes = list(shape = 16)),
    shape = guide_legend(override.aes = list(colour = "grey30"))
  ) +
  labs(
    x = NULL,
    y = "Bias vs reference (%)",
    title = "Accuracy - bias per analyte (QC & EQA; excl. DHEA, DHEAS)"
  ) +
  theme(
    aspect.ratio = 1,
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 6),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 5),
    axis.text.y = element_text(size = 5.2)
  )
p_eqa
```

## Figure 2 — manuscript figure

The four panels above are composed into the publication figure (**Figure 2**).

```{r chunk25-fig2-assemble}
#| label: fig-d4-fig2
#| fig-cap: "**MRMhub vs MassHunter — quantification.** Composite of the four panels above; reproduced as the manuscript's **Figure 2**. **(A)** Calibration-fit comparison, **(B)** concentration equivalence on identical areas, **(C)** full end-to-end agreement, **(D)** QC and EQA accuracy. Exported to `output/fig2_quantification_comparison.{pdf,png}` (180 mm)."
#| fig-width: 7.1
#| fig-height: 9.3
# Middle row (B, C) uses coord_equal(); give it enough height that the two square
# panels grow to fill the full width, so their L/R borders align with A and D.
# The bias panel is a square, 3-column grid as a standalone figure; inside the
# full-width composite it fills better as a wide 4-column grid without the forced
# square aspect, so override those two settings for Figure 2 only.
p_eqa_fig2 <- p_eqa +
  facet_wrap(vars(analyte_id), scales = "free_y", ncol = 4) +
  theme(aspect.ratio = NULL)
fig2 <- p_r2 /
  (p_qonly + p_full) /
  p_eqa_fig2 +
  plot_layout(heights = c(0.68, 1.3, 1.2)) +
  plot_annotation(tag_levels = "A")

save_fig(fig2, "fig2_quantification_comparison", height = 236)
fig2
```

# Summary

Across both pipeline stages, MRMhub-QUANT and Agilent MassHunter agree closely on
this steroid panel. At the **integration** stage, peak areas track the 1:1 line over
four orders of magnitude with comparable replicate precision, and the few larger
real-sample differences trace to legitimate integration-boundary choices on elevated
or co-eluting baselines rather than to a systematic bias (@fig-d4-fig1). At the
**quantification** stage, feeding MassHunter's own areas through the MRMhub pipeline
reproduces MassHunter's concentrations, isolating the post-processing as effectively
equivalent; the two independent end-to-end pipelines then show comparable accuracy
against the nominal LQC/HQC targets and the SKML EQA reference values (@fig-d4-fig2). Any residual differences are
attributable to peak integration rather than to the QUANT post-processing.

# References
