---
title: "Dataset 4 — Quantification & QC Report"
subtitle: "Steroid Assay"
author:
  - name: "Bo Burla"
    affiliation: "National University of Singapore"
  - name: "Guo Shou Teo"
  - name: "Hyungwon Choi"
date: "`r Sys.Date()`"
format:
  html:
    code-tools: true
    code-fold: true
    self-contained: true
    code-overflow: wrap
    toc: true
    number-sections: true
    df-print: paged
    lightbox: true
    fig-align: center
    fig-format: retina
    fig-dpi: 200
---

# Overview

This is an MRMhub-QUANT postprocessing report for Dataset 4, a **fully quantitative 15-analyte steroid panel** measured with external calibration (absolute quantification). The data are from an LC-MRM-MS analysis of steroids in human serum using the Steroid Panel LC-MS kit (IBL International, Hamburg, Germany; #30191875), following the Instructions for Use except that LC-MS was performed on an Agilent 6495A triple quadrupole instrument. Only the kit's Gradient 1 analysis was used, covering 15 analytes with their internal standards (66 MRM transitions monitored, including qualifiers, from which 64 features were integrated).

The batch comprised 15 system-suitability samples (SST, containing pure analyte and internal standards), blanks, calibrants (`Cal A`–`Cal F`), low and high quality controls (`LQC`, `HQC`) with assigned target concentrations, and — in place of study samples — 5 external quality assessment (EQA) samples (SKML ring-trial materials) [@jansen2014]. In both the MRMhub and MassHunter workflows, concentrations were obtained by external calibration using quadratic regression with `1/x` weighting for all analytes, and bias was determined for the QC samples against their assigned concentrations. Full source and acquisition settings, transitions, polarities and collision energies are listed in the INTEGRATOR result file (Zenodo, Dataset 4).

This report follows the MRMhub [external-calibration + QC recipe](https://slinghub.github.io/MRMhub/quant/articles/recipe-01-ext-calibration-qc.html). See `Dataset4_Comparison.qmd` for the comparison of MRMhub vs MassHunter across quantitative peak integration and postprocessing.

::: callout-note
## mrmhub version

Rendered with **mrmhub 0.9.9** (QUANT module of [MRMhub](https://github.com/SLINGhub/MRMhub), `main` branch): `pak::pak("SLINGhub/MRMhub@main")`.
:::

## Setup

Load the package set `mrmhub_enable_cli_color()` to renders the mrmhub's console feedback in colour in the HTML output.

```{r chunk2-loadpkg}
library(mrmhub)
mrmhub_enable_cli_color()
```

# Import

Two inputs are loaded into a fresh `MRMhubExperiment`. `import_data_mrmhub()` reads the INTEGRATOR long-format peak areas for the assay batch (blanks, calibrators, QCs and EQA samples). `import_metadata_msorganiser()` then attaches the msorganiser metadata workbook, which defines the analytes, their internal standards, the sample annotations (QC types, calibrator levels) and the calibration/target concentrations.

```{r chunk3-import}
mexp <- MRMhubExperiment(title = "Dataset 4 — Steroid Assay")
mexp <- import_data_mrmhub(
  mexp,
  path = "./data/dataset-4/Dataset4_MRMhub-INTEGRATOR_ASSAY.csv",
  import_metadata = TRUE
)
mexp <- import_metadata_msorganiser(
  mexp,
  path = "./data/dataset-4/Dataset4_Metadata.xlsx",
  excl_unmatched_analyses = TRUE, ignore_warnings = TRUE
)
```

# Calibration

`normalize_by_istd()` divides each analyte's peak area by that of its authentic stable-isotope-labelled internal standard. `calc_calibration_results()` then fits one external calibration curve per analyte from the `Cal A`–`Cal F` levels. A **quadratic** model with **`1/x` weighting** was used as the regression model for all analytes.

```{r chunk4-normalize-calibrate}
mexp <- normalize_by_istd(mexp)
mexp <- calc_calibration_results(
  mexp,
  fit_overwrite = TRUE,
  fit_model = "quadratic",
  fit_weighting = "1/x"
)
```

```{r}
#| label: fig-d4rep-calibration-curves
#| fig-cap: "**External calibration curves.** Quadratic, 1/x-weighted curves for the 15 steroid analytes (quantifier transitions)."
#| fig-width: 9
#| fig-height: 7
#| out-width: "100%"
#| fig.alt: "Calibration curves for the 15 steroid analytes."
# Curves are solid between the lowest and highest calibrator, and dotted where
# extrapolated beyond that calibrator range (automatic; not a range restriction).
plot_calibrationcurves(
  mexp,
  fit_overwrite = FALSE,
  fit_model = "quadratic",
  fit_weighting = "1/x",
  include_istd = FALSE,
  include_qualifier = FALSE,
  rows_page = 4,
  cols_page = 4,
  show_progress = FALSE
)
```

The metrics below summarise each quantifier's fit: **R²** (linearity/goodness of fit) and the estimated **LOD** and **LOQ** (limits of detection and quantification, in the calibrated concentration unit).

```{r}
#| label: tbl-d4rep-calibration-metrics
#| tbl-cap: "**Calibration metrics per analyte.**"
get_calibration_metrics(mexp, include_qualifier = FALSE, summary_table = TRUE)
```

# Quantification

`quantify_by_calibration()` inverts each fitted curve to convert the normalized intensities into absolute concentrations. Only quantifier transitions are used (`include_qualifier = FALSE`), and `ignore_failed_calibration = TRUE` skips any analyte whose curve failed to fit rather than aborting the whole run, so a single problematic analyte does not block the report. `fit_model` and `fit_weighting` can also be set per analyte in the feature metadata table.

```{r chunk7-quantify}
mexp <- quantify_by_calibration(
  mexp,
  fit_overwrite = FALSE,
  include_qualifier = FALSE,
  ignore_failed_calibration = TRUE,
  fit_model = "quadratic",
  fit_weighting = "1/x"
)
```

# QC bias & variability

The final check compares the measured QC concentrations against their known targets. `get_qc_bias_variability()` reports, per analyte and QC level, the mean measured concentration, the **bias** (percent deviation from target — *accuracy*) and the intra-batch **%CV** (*precision*), plus a flag for values falling outside the calibrated range. Restricting to `LQC`/`HQC` keeps the table to the samples that have assigned target concentrations.

```{r}
#| label: tbl-d4rep-qc-bias-variability
#| tbl-cap: "**QC bias and variability** for the LQC and HQC samples."
get_qc_bias_variability(mexp, qc_types = c("LQC", "HQC"), summary_table = TRUE)
```

# Export

The results are written in three complementary formats, each for a different use.

## Flat concentration table (CSV)

`save_dataset_csv()` writes a single flat table of concentrations, one row per sample — the simplest format to open in Excel or read into any tool. `add_qctype` keeps the QC-type column so calibrators, QCs and samples stay distinguishable.

```{r chunk9-export-csv}
save_dataset_csv(
  mexp,
  path = file.path("output", "Dataset4_report_conc.csv"),
  variable = "conc",
  add_qctype = TRUE
)
```

## Multi-sheet report (XLSX)

`save_report_xlsx()` writes a multi-sheet workbook bundling the concentrations *and* the supporting context (calibration metrics, feature QC metrics, raw/normalized intensities, sample and feature metadata) in one file for sharing and archiving.

```{r chunk10-export-xlsx}
save_report_xlsx(
  mexp,
  path = file.path("output", "Dataset4_report.xlsx"),
  filtered_variable = "conc",
  overwrite = TRUE
)
```

## Complete dataset object (RDS)

`save_dataset_rds()` writes the entire `MRMhubExperiment` — imported data, metadata, calibration fits and quantification results — to a single `.rds`. This is the format for **sharing a complete, self-contained dataset**: everything needed to reproduce, extend or audit the analysis travels in one file. It embeds an `rlang::hash()` fingerprint, so the file's integrity can be verified when it is read back. The file is a plain R object, so it opens in **any R session with `readRDS()`, even without mrmhub installed**; with mrmhub also loaded, the object's methods become available so the recipient can **re-process or inspect it** — picking up exactly where this report ends.

```{r chunk11-export-rds}
save_dataset_rds(mexp, "output/Dataset4_report_mexp.rds")
```

Reload it in a fresh R session with `read_dataset_rds()`, which verifies the embedded integrity hash on read (load `mrmhub` first so the object's methods are available):

```{r chunk12-rds-import}
#| eval: false
library(mrmhub)
mexp <- read_dataset_rds("output/Dataset4_report_mexp.rds")
```